---
title: "Interacting with Google Big Query to make a global map of fishing effort"
author: "Juan S. Mayorga"
date: "October 17, 2016"
output: html_notebook
---

This is a quick example of how to work with "Big data" in R through interacting with Google Big Query. We will use Global Fishing Watch data to show three different ways to access, analyze, and summarize dataset of millions and millions of rows.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

detach_package <- function(pkg, character.only = FALSE){
  if(!character.only)
  {
    pkg <- deparse(substitute(pkg))
  }
  search_item <- paste("package", pkg, sep = ":")
  while(search_item %in% search())
  {
    detach(search_item, unload = TRUE, character.only = TRUE)
  }
}

library(tidyverse)

```

# 1. Using the *bigrquery* package dplyr style:

```{r dplyr_style}
project <- "ucsb-gfw"

library(bigrquery)

fishing_connection <- src_bigquery(project, "Juan") # This function initiliazes a connection with a BQ dataset

effort_in_HS_by_countries <- fishing_connection %>% 
  tbl("clean_effort_with_VC_and_FAO_2015") %>%  #tbl connect to a specific table within dataset
  filter((is.na(eez) | eez == 'Antarctica') & measure_new_score >= 0.5) %>% 
  group_by(country) %>% 
  summarise(effort = sum(hours)) %>% 
  top_n(10, effort)
  
collect(effort_in_HS_by_countries) %>% 
  knitr::kable()

```

```{r joins_fail, eval = FALSE}
fishing_connection %>% 
  tbl("clean_effort_with_VC_and_FAO_2015") %>% 
  left_join(src_bigquery(project, "fao") %>% 
              tbl("cost_by_FAO_region_in_HS_2015"), by = "Fao_region")
```

Pros: 

  - Is basically dplyr on steroids
  - No need to learn another language
  - Easy to incorporate in your workflow
  
Cons: 

  - Not yet perfect: joins and some verbs and not fully compatible yet
  - Cannot query over multiple tables 

# 2. Using the *bigrquery* package SQL style:


```{r complex_queries_SQL_style, eval = FALSE}
sql <- "SELECT
  a.mmsi mmsi,
  b.country country,
  b.iso_code iso_code,
  b.is_EU is_EU,
  a.timestamp timestamp,
  a.lon lon,
  a.lat lat,
  a.speed speed,
  a.course course,
  a.heading heading,
  c.country_name eez,
  a.hours hours,
  a.distance_from_shore distance_from_shore,
  a.distance_from_port distance_from_port,
  a.seg_id seg_id,
  a.measure_new_score measure_new_score,
  a.FAO_Region FAO_region
FROM (
  SELECT
    *,
    INTEGER(IF( LENGTH(STRING(mmsi)) = 9,LEFT(STRING(mmsi),3), '0')) code
  FROM
    TABLE_DATE_RANGE([fao.], TIMESTAMP('2015-01-01'), TIMESTAMP('2015-12-31'))
  WHERE
    lat > -80
    AND lat <80
    AND lon > -180
    AND lon < 180
    AND mmsi > 100000000
    )a
LEFT JOIN (
  SELECT
    code,
    country_name country,
    iso3 iso_code,
    isEU is_EU
  FROM
    [vessel_lists.country_code])b
ON
  a.code = b.code
LEFT JOIN (
  SELECT
    *
  FROM
    [vessel_lists.eez_country_codes])c
ON
  a.eez = c.eez"

clean_names_2015 <- query_exec(sql,project, destination_table = "fao.test", max_pages = 10)
```

```{r mapping_resources}
library(maps)
library(rasterVis)

eez = rgdal::readOGR(dsn='../Clipperton/eez',layer='World_Maritime_Boundaries_v8') # get global EEZ shapefile
world = maps::map(database ="world", fill = TRUE, interior = FALSE, col = "gray",plot = FALSE) # get world shapefile
IDs <- sapply(strsplit(world$names, ":"), function(x) x[1]) # get polygon ID

bPols <- maptools::map2SpatialPolygons(world, IDs=IDs,
                             proj4string=sp::CRS("+proj=longlat +datum=WGS84")) # convert map object to a suitable sp object
```

```{r map_HS_effort_by_country_function}
map_effort_by_country <- function(year,country){
  
sql <-  paste("SELECT
INTEGER(floor(lat*2)) lat_bin,
INTEGER(floor(lon*2)) lon_bin,
SUM(hours) effort
FROM
  high_seas_cost_model.clean_effort_with_VC_and_FAO_",year,"
where 
  country = '",country,"' and
  measure_new_score >=.5
AND (eez IS NULL or eez = 'Antarctica')
group by 
lat_bin,
lon_bin", sep = "")

country_effort <- query_exec(sql,project)  
  
# Define map limits and grid size

max_lat = 90
min_lat = -90
max_lon = 180
min_lon = -180
cell_size  = 0.5
one_over_cellsize = 2

num_lats = (max_lat-min_lat)*one_over_cellsize+1
num_lons = (max_lon-min_lon)*one_over_cellsize+1

# create grid and fills it in 
grid <- matrix(nrow=num_lats,ncol=num_lons)

for (i in 1:nrow(country_effort)) {
  lat = country_effort[i, "lat_bin"]
  lon = country_effort[i, "lon_bin"]
  lat_index = lat-min_lat*one_over_cellsize
  lon_index = lon-min_lon*one_over_cellsize
  grid[num_lats - lat_index,lon_index] = country_effort[i, "effort"]
}

#creates rater from grid
r <- raster::raster(grid, 
                    xmn = min_lon,
                    xmx = max_lon, 
                    ymn = min_lat, 
                    ymx= max_lat, 
                    crs = NA, 
                    template = NULL) 

# define the theme: breaks, background, colorkey

my.at <- c(1,10,20,30,40,50,100,110,120,130,140,150,160,170,180,190,200,300,500)

myTheme <- rasterTheme(region=rev(colorspace::sequential_hcl(200, 
                                                             power=0.4,
                                                             h = 250, 
                                                             c = c(57,77), 
                                                             l = c(20,95))))
myTheme$panel.background$col = 'black' 

my.colorkey=list(at=my.at,
                 labels=list(at=my.at,
                             labels=my.at))
colorspace::choose_palette()  

# creates the map using levelplot
effort_map <- rasterVis::levelplot(r, 
                                   margin = FALSE, 
                                   par.settings = myTheme, 
                                   at = my.at, 
                                   main = paste(country,"Fishing effort in the high seas (hours) in",year,sep = " "),
                                   scales = list(draw = FALSE), 
                                   colorkey = list(colorkey = my.colorkey,
                                                 space ="bottom", 
                                                 width = 0.7))+
  layer(sp::sp.polygons(bPols, fill = 'gray', col = "gray"))

print(effort_map)

detach_package(rasterVis)
detach_package(raster)
detach_package(maps)}

map_effort_by_country(2015,"China")
```

Pros:

  - Can do more complicated queries and joins
  - query multiple table
  - Save query output to BQ (not as fast as BQ directly)
  - Can be functionalized
  
Cons: 

  - is SQL
  - Cognitive dissonance between languages
  - not the easiest thing to write SQL as string in R
  - not great to display code in dynamic documents


# 3. Using DBI, bigrquery, and RMarkdown chunks


```{r dbi_setup}
library(DBI)
BQ_connection <-  dbConnect(dbi_driver(),dataset = "", project = "ucsb-gfw")

country <- 'Mexico'
min_lat1 <- -20.4
max_lat1 <- 34.1
min_lon1 <- -133.6
max_lon1 <- -98.7

min_lat2 <- -20.4
max_lat2 <- 17.9
min_lon2 <- -99.8 
max_lon2 <- -75.2 
```


```{sql mexico_effort, connection=BQ_connection,  highlight=TRUE, output.var="mexico_ETP_effort"}
SELECT
  eez,
  mmsi,
  country,
  general_vessel_type,
  length,
  length_source,
  tonnage,
  tonnage_source,
  engine_power,
  sum(hours) fishing_hours,
  EXACT_COUNT_DISTINCT(DAYOFYEAR(timestamp)) fishing_days,
FROM
  [ucsb-gfw:high_seas_cost_model.clean_effort_with_VC_and_FAO_2015]
WHERE
  country = ?country
and measure_new_score >= .5 
 AND 
 ( ( lon > ?min_lon1 and lat > ?min_lat1 and lon < ?max_lon1 and lat < ?max_lat1 )
   OR 
   (lon > ?min_lon2 and lat > ?min_lat2 and lon < ?max_lon2 and lat < ?max_lat2 )
   )
group by 
  eez,
  mmsi,
  country,
  general_vessel_type,
  length,
  length_source,
  tonnage,
  tonnage_source,
  engine_power
having fishing_days > 1
```
  

```{sql mexico_effort_map, connection=BQ_connection,  highlight=TRUE, output.var="mexicos_binned_effort"}
SELECT
 INTEGER(FLOOR(lat*2)) lat_bin,
 INTEGER(FLOOR(lon*2)) lon_bin,
 SUM(hours) effort
FROM
 high_seas_cost_model.clean_effort_with_VC_and_FAO_2015
WHERE
 country = ?country
and measure_new_score >= .5 
 AND 
 ( ( lon > ?min_lon1 and lat > ?min_lat1 and lon < ?max_lon1 and lat < ?max_lat1 )
   OR 
   (lon > ?min_lon2 and lat > ?min_lat2 and lon < ?max_lon2 and lat < ?max_lat2 )
   )
GROUP BY
 lat_bin,
 lon_bin
```


```{r mexico_ETP_map}
library(maps)
library(rasterVis)

# define map's limits and grid size
max_lat = 35
min_lat = -10
max_lon = -75.2
min_lon = -133.6
cell_size  = 0.5
one_over_cellsize = 2

# initialyze a grid and fill it in 
num_lats = (max_lat-min_lat)*one_over_cellsize+1
num_lons = (max_lon-min_lon)*one_over_cellsize+1

grid <- matrix(nrow=num_lats,ncol=num_lons)

for (i in 1:nrow(mexicos_binned_effort)) {
  lat = mexicos_binned_effort[i, "lat_bin"]
  lon = mexicos_binned_effort[i, "lon_bin"]
  lat_index = lat-min_lat*one_over_cellsize
  lon_index = lon-min_lon*one_over_cellsize
  grid[num_lats - lat_index,lon_index] = mexicos_binned_effort[i, "effort"]
}

# create raster from grid
r <- raster(grid, 
            xmn = min_lon,
            xmx = max_lon, 
            ymn = min_lat, 
            ymx= max_lat, 
            crs = NA, 
            template = NULL) 

#define the theme: color key, background color, breaks 
my.at <- c(1,10,20,30,40,50,100,110,120,130,140,150,160,170,180,190,200,300,500)

myTheme <- rasterTheme(region=rev(colorspace::sequential_hcl(200, 
                                                             power=0.4, 
                                                             h = 240, 
                                                             c = c(57,77), 
                                                             l = c(20,95))),
                       layout.heights=list(xlab.key.padding=1))

myTheme$panel.background$col = 'white' 

my.colorkey=list(at=my.at,
                 labels=list(at=my.at,cex=.7,labels=my.at))

# get EEZ's 
library(mregions)
mr_names <- mregions::mr_names()
panama_eez_pacific <- mr_shp(name = "Panamanian Exclusive Economic Zone", read = TRUE)@polygons[[1]]@Polygons[1][[1]]
mexico_eez_pacific <- mr_shp(name = "Mexican Exclusive Economic Zone", read = TRUE)@polygons[[1]]@Polygons[4][[1]]
galapagos_eez <- mr_shp(name = "Ecuadorean Exclusive Economic Zone (Galapagos)", read = TRUE)
ecuador_eez_pacific <- mr_shp(name = "Ecuadorean Exclusive Economic Zone", read = TRUE)@polygons[[1]]@Polygons[4][[1]]
clipperton_eez <- mr_shp(name = "Clipperton Island Exclusive Economic Zone", read = TRUE)
colombia_eez_pacific <- mr_shp(name = "Colombian Exclusive Economic Zone", read = TRUE)@polygons[[1]]@Polygons[1][[1]]
costa_rica_eez_pacific <- mr_shp(name = "Costa Rican Exclusive Economic Zone", read = TRUE)@polygons[[1]]@Polygons[1][[1]]
guatemalan_eez_pacific <- mr_shp(name = "Guatemalan Exclusive Economic Zone", read = TRUE)@polygons[[1]]@Polygons[1][[1]]
nicaraguan_eez_pacific <- mr_shp(name = "Nicaraguan Exclusive Economic Zone", read = TRUE)@polygons[[1]]@Polygons[1][[1]]
salvadorean_eez_pacific <- mr_shp(name = "El Salvador Exclusive Economic Zone", read = TRUE)@polygons[[1]]@Polygons[1][[1]]

# make shapefile of proposed MPA

revilla_coords <- read_csv('../Revillagigedo/revilla_puntos.csv')
revilla_p <- Polygon(revilla_coords, hole = TRUE) 
revilla_ps <- Polygons(list(revilla_p),1)
revilla_sps <-  SpatialPolygons(list(revilla_ps), proj4string=sp::CRS("+proj=longlat +datum=WGS84"))

# make the map with levelplot

ETP_effort_map <- levelplot(r,
          margin = FALSE, 
          par.settings = myTheme, 
          at = my.at, 
          main = "Mexico's Fishing effort in the Eastern Tropical Pacific",
          scales=list(draw=FALSE), 
          colorkey=list(colorkey = my.colorkey,space="bottom", width = 0.7, title = "hours"), 
          xlab=list(label="Fishing hours", cex=.8),
          cexRow=0.8, cexCol=0.8)+
  layer(sp::sp.polygons(bPols, fill = 'gray', col = "gray"))+
  layer(sp::sp.polygons(panama_eez_pacific))+
  layer(sp::sp.polygons(ecuador_eez_pacific))+
  layer(sp::sp.polygons(mexico_eez_pacific))+
  layer(sp::sp.polygons(galapagos_eez))+
  layer(sp::sp.polygons(colombia_eez_pacific))+
  layer(sp::sp.polygons(clipperton_eez))+
  layer(sp::sp.polygons(costa_rica_eez_pacific))+
  layer(sp::sp.polygons(guatemalan_eez_pacific))+
  layer(sp::sp.polygons(nicaraguan_eez_pacific))+
  layer(sp::sp.polygons(salvadorean_eez_pacific))+
  layer(sp::sp.polygons(revilla_sps, col = 'red'))

print(ETP_effort_map)

detach_package(rasterVis)
detach_package(raster)
```

