---
title: "Learning SF"
output: html_notebook
---

This is a notebook for learning the `sf` spatial analysis package, to help better deal with spatial files when you get them, specifically in support of skynet

Specifically, your goal is to be able to more seamlesslely move back and forth from UTM stuff in from spatialdeltaGLM to lat-lons when you get them, and to be able to do a better job making maps and other joyous things. 

```{r}
library(tidyverse)
library(sf)
library(maps)
library(ggmap)

```

`st_` refers to spatial and temporal 

`sf` is the data table containing the spatial stuff

`sfc` is the list column of geometries for a given feature

`sfg` makes up the components of each list component in `sfc`

```{r}

nc <- st_read(system.file("shape/nc.shp", package="sf"))


class(nc)

head(nc, n = 6)
```
So suppose that I want to look more at the actual geometries of each area, can access as a column, or use st_geometry as long as the object is of class `sf`

```{r}

nc_geom <- st_geometry(nc)

```

