---
title: "Estimating Fish Abundance from the Behavior of Fishing Fleets"
authors: "Dan Ovando, Chris Costello, Steve Gaines, Ray Hilborn, Jim Thorson"
output:
  bookdown::pdf_document2: null
  bookdown::html_document2: null
  html_document: default
linkcolor: blue
bibliography: dissertation.bib
biblio-style: apalike
---

```{r, include=F}
knitr::opts_chunk$set(message = F, warning = F, echo = F,cache = F)
```


```{r, include=FALSE}
library(brms)
library(sf)
library(viridis)
library(scales)
library(rstan)
library(patchwork)
library(hrbrthemes)
library(tidyverse)
extrafont::loadfonts()


fig_theme <- theme_ipsum(base_size = 14, axis_title_size = 18)

theme_set(fig_theme)

map_theme <-   theme_get() +theme(legend.key.height = unit(1.5,"cm"),
                                  axis.text.x = element_blank(),
                                  axis.text.y = element_blank(),
                                  plot.margin = unit(c(0,0,0,0),"cm"),
                                  legend.box.margin =  ggplot2::margin(0,0,0,0))

functions <- list.files(here::here("functions"))

walk(functions, ~ here::here("functions", .x) %>% source()) # load local functions

run_name <- 'd1.1'

run_dir <- here::here('results',run_name)

load(file = "/Users/danovando/PhD/skynet/results/d1.1/skynet-plots.Rdata")

```

## Literature notes

see @vanPutten2012 for references supporting fish abundance as a driver of economic activity

@Gillis2012 for an alternative economic model to the ideal free distribution. Also has great citations for behavior tracking abundance for snow crabs, as well as failures for other fisheries


## Executive Summary

The goal of this study is to determine the ability of Global Fishing Watch (GFW) derived data on fishing activity to predict fish abundance as estimated from fishery independent research surveys.

We do this by testing the ability of structural and machine learning models to use GFW data to make accurate out-of-sample predictions.

Current results suggest that structural approaches are not capable of this task, but the machine learning approaches show potential.

The model can predict absolute and relative spatial abundance with some accuracy. The model can also capture overall temporal trends, but struggles to perform for time periods excluded from the model training.

### Next Steps

  - Fitting the model using only GFW derived data (no environmental covariates)
      - might degrade in-sample performance but improve out-of-sample performance

  - Fitting the model using only unfished species
      - If we are truly capturing a "fishing signal", the the performance of a model fit to unfished species should degrade substantially (though we would still expect it to work somewhat due to covariance between fished and unfished species)

  - Getting our hands on a much more "out of sample" dataset: North Sea, Mediterranean, West Africa?

  - Determining a method for benchmarking performance: How good is good enough?
      - Compare performance of Catch-MSY/surplus production model estimates of abundance trends?
      - ....?

  - Do one more pass at a "keep it simple stupid" approach: Get catch data for the survey regions for all the species in the surveys, take total catch divided by total effort from GFW, see how that index of multi-species abundance compares to survey based estimates.



## Introduction

Successful fisheries management rests in part on the ability to provide accurate and timely assessments of the status (generally in the form of biomass levels and/or fishing mortality rates relative to some reference point) of fish stocks. Fisheries science has developed an expansive and often effective toolbox for providing this knowledge, but the data-intensive nature of many of these tools has prevented their use in all but the most knowledge and resource rich parts of the world. In recent years, this problem has led to an rapid expansion of "data-limited assessments" (DLAs), that seek to provide stock status estimates using fewer data (but more assumptions) for fisheries that do not have for formal stock assessments in place [such as those encompassed by the RAM Legacy Stock Assessment Database, @Ricard2012]. While length-based DLAs are commonly used at the more local level [e.g. @Hordyk2014; @Rudd2017], at larger spatial scales catch-based methods, that try to explain stock status as a function of trends in the amount of fish caught from a population by fishing pressure, have become the standard method [e.g. @Costello2012a; @Costello2016; @Rosenberg2017].

This prevalence of catch-based methods is based largely out of necessity rather than performance; catch statistics, such as those provided by the Food and Agriculture Organization of the United Nations (FAO; @FAO2018 citation), have been to date the only globally available source of fishery statistics. While these methods have been shown to be effective in some circumstances [@Rosenberg2017, @Andersona], catch statistics alone can be misleading in understanding stock status [@Pauly2013]. The need for expanded tools to rapidly understand and manage data-limited fisheries is growing increasingly critical as populations grow and the climate changes. Global Fishing Watch, presented by @Kroodsma2018, presents a new global database of fishing effort, updated in near real time XX. Can these new data be used to improve our ability to understand the status of fisheries around the world?

Why might we think that data on the dynamics of fishing effort might be useful for fisheries stock assessment? Regardless of their scale, from large industrial operations to small artisanal activity,  fisheries share a common underlying incentive structure: fishermen desire some utility derived from capturing fish (e.g. some combination of food, income, and cultural activity) and tune their fishing activities in order to try and maximize that utility, subject to the constraints of the world. As time goes on, these fishing actions affect fish stocks, for example through reductions in abundance or mean length, causing the behavior of the fishermen to be updated. In short then, fishing and fish are part of a dynamic system, in which the behavior of each affects the behavior the other.

These dynamic links between fishing fleets and fish populations are a critical part of fisheries management. @Gordon1954 predicted that in the absence of property rights or access restrictions, these bio-economic dynamics would result in the fishery reaching an open-access equilibrium at which net profits in the fishery are driven to zero, often resulting in biological overfishing of the stock. This thinking was central to the bleak predictions of the "tragedy of the commons" [@Hardin1968]. While @Ostrom1990 demonstrated that the evolution of natural resource systems such as fisheries are driven by a more complete and complex set of drivers than pure profitability, the critical link between the dynamics of fish populations and fishing communities remains in the form of social-ecological systems [@Ostrom2009].

Understanding the dynamics of these social-ecological systems is critical to effective fisheries management. Bio-economic theory and empirical evidence helps us design and implement policies that best achieve societal objectives, by allowing us to model what the potential impacts of policy choices may be. This process has underpinned the recent expansion in the use of rights-based fisheries management [e.g. @Costello2008; @Costello2016; @Deacon2007; @Grainger2016; @Kaffine2008; @Wilen2012; @Cancino2007; @Grafton1996; @Squires2016], and increasingly is used in the management-strategy-evaluation [MSE, @Butterworth2007; @Punt2016a] process. @Nielsen2017 provides a thorough review of models currently utilizing bio-economic modeling to model and guide the policy implementation process.

These works demonstrate a rich history of thinking about how fishing fleets respond to incentives and population dynamics. However, the vast majority of the literature in this field is focused on predicting fishing effort as a function of fish populations; very little research has reversed this question and asked, what do the dynamics of fishing effort suggest about the state of fish populations? If fisheries are indeed a coupled bio-economic system, then just as we believe the dynamics of fishing effort should be predictable from fish abundance, fish abundance should be predictable from fishing effort [@Sugihara2012]. The reason for this omission is likely driven by the realities of fisheries management. Bio-economic modeling is almost exclusively used in the MSE phase of fisheries management, rarely in the stock assessment phase itself. This is partly due simply to the history of stock assessment as a biological science [@Smith1994], and partly due to data realities. Historically, most fisheries with say accurate and complete records of fishing effort would also have records of more directly useful (from the perspective of traditional stock assessment) indices such as catches and catch-per-unit-effort. In these cases the catch data, combined with catch-per-unit-effort (CPUE) data, can provide a clearer signal of the state of a fish stock (with the catch data providing information on the scale of abundance and CPUE data on trends) than the effort data.

<!-- Many fisheries management policies are based around first achieving a biological objective (e.g. not overfishing the stock), and policies are then evaluated in terms of alternative strategies for achieving that objective.  -->

While effort dynamics have not been historically used to understand fish populations, it is reasonable to believe that given sufficient data, it should be possible to infer something about the state of a fish population as a function of the behavior of fishing fleets. This idea is in many ways analogous to earlier research linking the behavior of sea birds and the location of their prey [citation]. While this idea is grounded in sound theory, the actual form of link between fishing effort and fish abundance is far from clear. High levels of fishing effort could reflect high abundance of fish in the earlier years of a fishing ground, or could reflect an overfished but easily accessible region. We hypothesize that all else equal, fishermen would like to maximize their utility from fishing, but the varied and complex nature of these individual utility functions, combined with the shifting and uncertain nature of the natural world, make the structure of the link between fishing behavior and fish abundance a complicated question.

Empirical evidence and theory suggest that a) there is a link between effort and abundance but that b) the dynamics of those linkages can be complicated. This project tests the ability of different models to untangle these dynamics and use effort data from Global Fishing Watch to make accurate predictions of fish abundance. We do this by pairing these effort data with fishery-independent research surveys compiled by the [FishData](https://github.com/James-Thorson/FishData) package in R [@RCoreTeam2018] to ask, can the GFW derived effort data be used to predict the abundance of fish?

We break this assessment into a series of discrete phases.

  - We assess the relationships between total effort and biomass indices in order to see if, simply put, more more fishing is associated with more or less fish

  - We assess whether total effort can be paired with regional catch data to create an index of catch-per-unit-effort which in turn provides a reasonable index of fish abundance

  - We consider the ability of a series of structural economic models, based on bioeconomic theory, to predict fish abundance as a function of fishing behavior

  - We compare the predictive performance of the structural models to a suite of machine-learning models that utilize the same GFW data, and available covariates if desired, to predict fish abundance

  - Lastly, we consider the predictive value of GFW derived information relative to other globally available indices (e.g. sea surface temperature and chlorophyll)

## Methods

### Picking Models for Prediction

Assuming that methods such as the research surveys used in this study represent our best estimates of true stock status (i.e. that metrics such as effort do not somehow provide better estimates of status than the surveys, which seems a safe assumption), the ability of effort to predict abundance if we have reasonable belief that it will work in places that have effort data but do not have research surveys. For example, we could envision using this proposed effort-based model in locations that are covered by Global Fishing Watch but not by research surveys (which would represent most of the globe), or in between survey years for non-annual research surveys (e.g. those in the Aleutian Islands or the Gulf of Alaska).

We need then some method for judging which candidate models are likely to be the best at out-of-sample prediction. We did this by following the framework laid out in @Kuhn2013. We first took our merged data base and created a series of training and testing splits. These training splits were then used for all model fitting and judging. The testing splits were held aside until model selection based on the training data was complete, at which time we tested the ability of the tuned models to predict the density of fishes reported in the testing data. This specific slow is critical to selecting the model with highest chance of providing good out-of-sample prediction. Even if the testing splits are not used in the fitting directly, repeatedly fitting the models to the training data and inspecting the performance against the testing data introduces an element of bias where the selected model may be the one that happens to do best for that specific test-training split, and as such is not truly independently tuned for out-of-sample prediction.

We created a large number of candidate test and training splits, in order to compare the sensitivity of the model selection process during the training phase to the specific splitting.

  - `random`
    - The data were split using stratified random sampling from within the survey regions (to ensure that all regions were proportionally covered)
  - `spatial`
    - We used a spatial variogram to determine the approximate distance (in km) at which the spatial correlations in densities in the survey data plateaued. We then created the training data by selecting points separated by the amount of distance specified by the variogram, and the testing data for the points in between. We did this to try and prevent the model from exploiting the spatial correlation in the data to appear better at "out of sample" prediction than it in fact is
  - `california`
    - We split the data into a training set of data off of the coasts of Washing and Oregon state, and used data off of California for the testing split. This split helps us evaluate whether a model fit to one region can be extrapolated to another region, albeit a roughly similar one in some areas.

For the sake of clearer results, we only presen our diagnostics of the model fitting to the training data using the `random` split, but we return to the `spatial` and `california` splits when we confront the selected models to the testing splits.


## Methods

### Data

GFW provides our data on the amount and location of fishing effort, along with available observed or estiamted covariates such as vessel size, distance from shore/port, and engine power. Estimates of fish abundance (measured as density or biomass) in space and time were obtained from their relevant surveys through the `FishData` package. The following two data sections provide summaries of the data as well as descriptions of relevant data processing steps taken.

#### Trawl Survey Data

`FishData` provides access to numerous fishery independent research surveys throughout the world. We use the bottom trawl surveys conducted along the west coast of North America by the National Marine Fisheries Service  (\@ref(fig:knot-map))

  - [Eastern Bering Sea Bottom Trawl Survey](https://www.afsc.noaa.gov/RACE/groundfish/bottom%20trawl%20surveys.php) (ebsbts)
  - [Aleutian Islands Bottom Trawl Survey](https://www.afsc.noaa.gov/RACE/groundfish/bottom%20trawl%20surveys.php) (aibts)
  - [Gulf of Alaska Bottom Trawl Survey](https://www.afsc.noaa.gov/RACE/groundfish/bottom%20trawl%20surveys.php) (goabts)
  - [West Coast Groundfish Bottom Trawl Survey](https://www.nwfsc.noaa.gov/research/divisions/fram/groundfish/bottom_trawl.cfm) (wcgbts)
  <!-- - West Coast Groundfish Hook and Line Survey -->

```{r knot-map, fig.cap='Spatial coverage of fishery independent research surveys used in this study'}

knot_plot

```

Each of the surveys contains data on a wide variety of different species, including highly abundance fished species such as arrowtooth flounder (*Atheresthes stomias*) and Alaska pollock (*Gadus chalcogrammus*), as well as unfished species such as miscellaneous sea anemones (order *Actiniaria*). The selected surveys utilize bottom trawl gear, and as such primarily contain bottom-associated species (\@ref(fig:fish-plot)). Surveys are conducted in summer months (July-August for the Alaska surveys and May-October for the West Coast Bottom Trawl Survey).

```{r fish-plot, fig.cap = "Number of positive encounters for the top-10 most observed species in each research survey"}

fish_plot

```

Survey data are provided by `FishData` in their "raw" form (biomass by species per unit of survey effort at a given sampling event). The data therefore require standardization to account for differences in vessel characteristics, spatio-temporal correlation structures, and the presence of zeros in the haul data. This standardization was performed using the [`VAST`](https://github.com/James-Thorson/VAST) package [@Thorson2016a], which implements a spatial delta-generalized linear mixed model to provide a standardized spatio-temporal index of abundance for each species in the data. While versions of `VAST` allow for accounting of both within and across species correlations, we chose to run the standardization process seperately for each species for the sake of convergence time (tests of this choice on smaller subsets of the data indicate the differences between the two approaches are not substantial).

The result of VAST is a network of "knots" that define polygons of equal density for each species, where the density of each species is measured in units of tons/km^2^ XX (Fig.\@ref(fig:fish-map)).

```{r fish-map, fig.cap = "VAST estimates of density (tons/km^2^) of species observed in bottom trawl surveys for the Alaska region"}

  fish_map_plot

```


Surveyed species vary substantially in their economic importance. We mark species encountered in the surveys as "fished" if their name, or a synonym for their name identified through the `taxize` package, was found within the global catch records of the Food and Agriculture Organization of the United Nations [@FAO2018]. For each species, we also obtained price estimates using the data provided by @Melnychuk2016. Together, these data provide estimated density for fished species encountered by the US west coast bottom trawl survey program over space and time, along with the associated value of these species. This allows us to examine both the density of species, and the "revenue density" available for fishing in different locations (Fig.\@ref(fig:price-plot)).

```{r price-plot, fig.cap = "Estimated fishing revenue density ($/km^2)"}
revenue_map_plot
```

#### Fishing Effort Data

Fishing effort data were obtained using the `bigrquery` package in R from Global Fishing Watch. Data were aggregated to the resolution of year and nearest 0.2 degree latitude and longitude, and for each vessel at this a given location we calculated the total fishing hours spent there, average distance from shore, average distance from port, and whether that location is inside an MPA (and if so whether the MPA was no take or restricted use). We also collected relevant data for that vessel such as its engine power, length, tonnage, and vessel type (trawler, purse-seine, fixed-gear, etc). Together, these data provide fishing effort-related data covering the regions surveyed in our fishery-independent data (\@ref(fig:gfw-map)).

Global Fishing Watch uses a neural-net to classify observed behavior as fishing or not fishing. From there, we filtered out the classified fishing behavior to entries that were more than 500m offshore, were moving faster than 0.01 XX check units and slower than 20 XX check units, to remvoe likely erenoeous fishing entries.

```{r gfw-map, fig.cap="Total hours of fishing activity reported by Global Fishing Watch within the study regions"}
gfw_plot
```

#### Environmental Covariates

We augmented the effort and abundance data with globally accessible environmental covariates of

  - Chlorophyll
  - Sea surface temperature
  - Bathymetry

All data were obtained from NOAA ERDDAP portal (https://coastwatch.pfeg.noaa.gov/erddap/index.html), and aggregated as needed to match the resolution of the GFW data (annual and 0.2 degree lat/long resolution). Other environmental data were explored (e.g. wave and wind), but did not have sufficient near-shore coverage for inclusion in the model.

#### Creating Merged Database

Having pulled together data on fish abundance, fishing effort, and environmental covariates, we then merged these data together into a comprehensive database. Effort data and environmental data were first merged together by matching time and location (as measured by latitude and longitude). The combined data were then clipped to only include observations that fall within the boundaries of the polygons defined by each trawl survey (Fig.\@ref(fig:knot-map)). From there, we snapped each effort-environment observation to the nearest (in terms of latitude and longitude distance) knot of fish abundance as defined by the `VAST` standaridization process. Since the survey data are generally at a courser resolution than the effort data, this means that multiple effort observations are often associated with any one knot at any one time.

We then performed a series of filtering steps on this merged database. Since all surveys are geared towards bottom dwelling species, only bottom-associated gears are included from the effort data.  In this case that means vessels identified by Global Fishing Watch as trawlers, pots and traps, and set gears (longline and gillnet). We also only included species that were observed 10 or more times during each year of the survey to improve model convergence. To account for potential seasonal shifts in abundance, we also filtered the data down to months in which the relevant research trawl surveys were conducted.


### Candidate Predictive Models

We evaluated three classes of candidate models for linking fishing effort to fish abundance:

  1. Linear models
    - These simply link abundance to effort through linear models.
  2. Structural economic models
    - These assume a non-linear functional form to the allocation of fishing effort, and tune the models to available data conditional on these structural assumptions.
  3. Machine learning models
    - These models make no explicit structural assumptions, but rather
    find the combination of predictor variables that maximize the out-of-sample
    predictive power of the model

The choice of evaluating both structural and machine learning models is important to discuss for a moment. Substantial amount of empirical evidence and bio-economic theory exists hypothesizing how fishing effort and fish abundance might be related, from simple ideal-free distributions to complex agent-based models. These structural models have the advantage of interpretability, but leave us open to errors in model specification. In contrast, machine learning models lose interpretability but are less sensitive to specification errors. While different in their mechanics, all the candidate machine learning models are black-box models whose sole objective is to maximize the predictive power of the model (as defined by the user). The user specifies some model options, but the model decides which data are important and how those data relate to each other. This allows these algorithms to fit highly non-linear models (if the data demand it), without the need to specify an exact statistical or structural form for how variables such as costs, safety, and fish abundance interact to affect fishing behavior.

As a result, machine learning models can serve as an effective benchmark for the best possible ability of GFW data to predict fish abundance. The disadvantage is that, while new techniques are emerging for interpreting machine learning model fits, they are inherently black boxes and as such do not permit us to really interpret the meaning of specific coefficients. The lack of a structural theory behind the model may also hamper the ability of these models to predict radically out of sample data (e.g. a machine learning model trained in Alaska may perform terribly in Africa). By running both, we can compare the machine learning and structural approach and see how much the interpretability of the structural model "costs" us in terms of predictive power, relative to the benchmark of the machine learning model.

### Linear Models

We include the linear models purely for data exploration (though if they happen to work well they could be used). The linear models include simple linear regressions between metrics of effort and metrics of abundance (e.g. total hours against total biomass). We also consider CPUE trends as a class of linear model (since it is just a linear transformation of the effort data), where we now ask is CPUE proportional to abundance. We tested one slightly more involved linear model, of the forms

$$(E_{y,k}) \sim Gamma(cost_{y,k} + \hat{B_{y,k}}, shape, scale)$$

$$ \hat{B_{y,k}} ~ normal(\bar{B_{survey}}, sigma_{survey})$$

In order for $\hat{B}$ to in fact be representative of fish biomass at a location, the assumption is that all of the other attributes that affect the decision of how much effort to allocate at a given site are captured by the estimated cost coefficient, which we assume is a linear function of the distance of a knot *k* from port, the distance from shore, the depth at that location, and the mean vessel size used at that observation. We used a Bayesian hierarchical model implemented through the `rstanarm` package to estimate this linear model. We then extracted the estimated *space* coefficients, and compared them to the fish biomass estimated at that location by the relevant fishery independent survey.

### Structural Models

Our structural models are constructed in the same manner as @Miller2016. The key of this model is the assumption that for a given spatio-temperal resolution, fishermen distribute themselves such that marginal profits are equal in space.

Following @Miller2016, we consider marginal profits $\Pi$ per unit effort as being

$$\Pi_{y,k} = pqB_{y,k}e^{-qE_{y,k}} - c_{y,k}$$

where for year *y* at knot *k*, *p* is price [drawn from @Melnychuk2016], *B* is our index of abundance (from the trawl surveys and VAST), *q* is catchability, *E* is effort (supplied by Global Fishing Watch), and *c* are variable costs).This leaves *q* and *c* as the unknowns in this equation.

@Miller2016 was primarily interested in estimating quota price aspects of *c*, taking as data *p*, *CPUE*, *E*, and other components of *c* (fuel, labor, ice, etc.). We are instead interested in estimating CPUE as a function of other variable, and so we can rearrange this equation to construct a model of the form

$$log(B_{y,k}) \sim normal(\frac{\Pi_{y,k} + c_{y,k}}{pqe^{-qE_{y,k}}}, \sigma_B)$$

Similar to @Miller2016 we assume for now that $\Pi_{y,k}$ is zero, though this is clearly not accurate given that many of the fisheries sampled by this model are highly regulated and in some cases rationalized (however, changing $\Pi_{y,k}$ to positive values had little effect on the fit of the model during trial runs). That leaves *q* and *c* as unknown parameters. While we do not have the high resolution logbook data available to @Miller2016, we could certainly obtain data on fuel and labor prices for this model. However, at this time, we simply assume that $c_{y,k}$ is a linear function of the distance of a knot *k* from port, the distance from shore, the depth at that location, and the mean vessel size used at that observation. We fit this model, estimating *q* and the cost coefficients and $\sigma_B$ using maximum likelihood with the Laplace approximation implemented with Template Model Builder (TMB, citation) in R.

This form of the model assumes the goal is to estimate *B* directly. Use of this model for prediction in new locations would require assuming that the fitted *q* and *cost* values are applicable to a new location. An alternative approach is to estimate a vector of latent variables $\hat{B}$ that, together with the *cost* and *q* estimates explain the observed effort distribution.

$$(E_{y,k}) \sim Gamma(\frac{1}{q}log(\frac{pq\hat{B_{y,k}}}{c_{y,k} + \Pi_{y,k}}), shape, scale)$$

$$ \hat{B_{y,k}} ~ normal(\bar{B_{survey}}, sigma_{survey})$$

This form of the model can be custom fit to any new region, but requires the assumption that all of the non-biomass related reasons for fishing at a given site are captured by the fitted *cost* coefficients, leaving the "biomass" effect to be captured by the latent variables.  If there are other site specific factors that affect the amount of fishing effort though, these factors will get soaked up by $\hat{B_{y,k}}$, confounding the interpretation of these fitted latent values as biomass indicators.

### Machine Learning

We implemented four machine learning algorithms:
  - random forests (implemented through the `ranger` package in R)
  - generalized boosted regression modeling (`gbm`)
  - boosted multivariate adaptive regression splines (boosted `bmars`)
  - multivariate adaptive regression splines (`mars`)

An important feature across all the machine learning approaches is that they all adaptively push back against predictive overfitting. Within the training data split, the machine learning approaches then split the training data into numerous new testing and training splits (typically now called assessment and analysis splits). The coefficients of the model are then in part fit by repeatedly searching subsets of parameters that minimize the predictive error of the model trained on the analysis split. Tuning parameters can be selected by comparison of predictive error of fitted models applied to the assessment splits. This process is repeated thousands of times the algorithms search for coefficients that while fitted on one set of data still provide reasonable predictive power on a held out set of data.

A random forest works by fitting a series of regression trees. Each regression tree takes a sub-sample of the training data, and a sub sample of the independent variables provided for model fitting. The algorithm then determines the variable and variable split (e.g. vessel size and vessel size above 30ft) that provides the greatest explanatory power for the sampled data, and creates that as the first node. The next two nodes are selected in the same process, and so on and so forth, down to a specified tree depth tuned through the `caret` package. Each tree provides a high-variance, low bias estimator of densities. The random forest then averages over hundreds of trees to reduce this variance and provide an improved estimate of density as a function of provided covariates. The advantage of this approach is that it makes no assumptions about error distributions or linearity of parameters, and the process of randomly sampling both data and predictors actively pushes back against overfitting [@Breiman2001].

Despite the starkly different name, a GBM is more or less a modification of a random forest that helps the model target and improve the fit of parts of the data that the model struggles with. It does this by for each split, calculating the residuals, and then adapting the model fit to target parts of the data with large residuals [citation]

Multivariate adaptive regression splines (MARS) models exploit similar assessment-analysis tools as random forest, but rather than working by splitting the data into a series of discrete bins that form a tree, the model breaks the data and variables into an ensemble of linear regressions. For example, consider a process *f* that can be modeled by two linear models, y when x is 1:10 y ~ f(0.1 * x), and when x is 11:20 y ~ f(0.5 * x). A properly tuned MARS model will search through the data, notice the split, and fit two different linear regression to each component of the data. Similarly to the random forest model, the MARS model considers subsets of available variables and possible levels of interaction among these variables. The "boosted" version of this model targets hard-to-fit parts of the model in the same was as the GBM model.

Unlike more classic fitting procedures, like say considering a GAM vs a GLM, there is little *a priori* reason to consider one form of these machine learning models over another. They all have been shown to work well under different circumstances, and so the selection process often simply comes down to selecting some reasonable candidates, fitting them, and then selecting the best model from the fits to the training data based on the user's criteria.

## Results

### Linear Models

Before heading down the statistical rabbit hole, we can simply examine how well linear transformations of effort predict abundance. This has an intuitive aspect to it; we can hypothesize that the reason that more fishing occurs in the challenging waters off Alaska than Santa Barbara is that there are higher volumes of valuable fish in that area. However, we could also imagine a scenario where fishing effort is concentrated in overfished but inexpensive grounds, leaving higher fish abundance in more remote areas that are not economical to fish.

Looking at the effort and abundance indices, we see some evidence of a "more fishing where there's more fish" hypothesis. Across each of the survey regions, aggregating up to a 200 km^2^ area, there is a positive correlation between the total hours of applicable fishing observed by GFW in that area and the total estimated revenues available in that area (estimated by the sum of the density per knot times the area of that knot times the price of each price of each species in that knot within the area encompassed by the 200 km^2^ box). However, the relationship is far from clear, with substantial variation around the mean slope for each region. In addition, we see if all one knew was the total amount of fishing hours, the magnitude of the fishing opportunity that those hours might correspond to varies substantially. (Fig.\@ref(fig:effort-v-fish)).

```{r effort-v-fish, fig.cap = "Total fishing hours plotted against fish revenue density. Available revenue is calculated as the sum of the densities of each species times their respective ex-vessel price. Each point represents a 200km^2^ area"}

e_v_ed_plot

```

This coarse data analysis suggests that there may be a relationship between total fishing effort and the value of the fishing opportunity within a region, but certainly not a clear enough relationship to serve as a reliable predictor of fishable biomass. That effort levels alone are not clearly informative is not very surprising. What though do we learn by pairing effort data with locally available catch data to construct a CPUE index? CPUE can, under the appropriate circumstances, serve as an index of relative abundance. To create a GFW derived CPUE index, we pulled catch data for the relevant regions and species from three different databases: the RAM Legacy Stock Assessment Databse [@Ricard2012], the NMFS commercial landings database [citation], and the FAO's capture production database [citation]. Pairing these catch data with the the GFW effort data gives us a timeseries of aggregate CPUE for a given region. We then compared these CPUE trends to trends in biomass provided by the RAM database and from the processed trawl survey used throughout this study. We see no clear agreement between these two metrics, and if we take the RAM and survey derived indices of abundance as true, the GFW derived CPUE index is in some cases provides an exactly opposite index of abundance from the true value (Fig.\@ref(fig:gfw-cpue))

```{r gfw-cpue, fig.cap="GFW derived CPUE (yellow lines) and assessments of abundance (blue) for the US West Coast and Eastern Bering Sea regions"}

wc_cpue_plot + ebsbts_cpue_plot
```

Turning to the latent variables approach for the linear models, We find no correspondence between the estimated space effects $\hat{B}$ and the fish biomasses estimated from the trawl surveys. This does not mean that such an approach might not work given sufficient data, but with the available covariates either too many other important factors besides biomass are being absorbed into the $\hat{B}$ coefficients, or the survey biomass is a small component of the decision making process for a given fishing location (Fig.\@ref(fig:linear-latent-plot)).

```{r linear-latent-plot, fig.cap="Scaled latent biomass coefficients plotted against paired scaled biomass estimates. Red dashed line indicates a 1:1 fit, solid blue line represents a linear model of the two axes"}

linear_latent_plot

```

### Structural Models

Since raw effort and effort derived CPUE indices do not appear to be valid methods for estimating abundance, we now turn to more detailed modeling approaches to utilize effort data to predict abundance. Our structural modeling approach follows a standard bio-economic framework, as laid out in @Miller2016. @Miller2016 used a structural modeling approach in part to estimate the quota prices in the US West Coast groundfish trawl fishery individual fishing quota (IFQ) program, using data on logbook reported CPUE, prices, and variable and fixed costs (labor, fuel, etc.). Making the assumption that for an appropriate fleet unit and time period marginal profits are equal in space, @Miller2016 then estimates quota prices for different species that, given their other data, rationalize the observed distribution of effort in the fishery. The biomass-predicting structural model shows limited predictive ability within the training set, with R^2^ within the training data less than 0.1.

```{r strct-ovp-plot, fig.cap= "Observed vs predicted biomass (A) and R2 values (B) for the fitted structural models across different evaluated data splits"}

strct_ovp_plot

```

We can also invert this idea in the same manner as we did with the linear model, and rather than estimating cost coefficients that explain the observed fish abundance, given observed efforts, estimate cost coefficients and latent spatial parameters representing abundance that explain the observed effort. We estimated this model using a Bayesian hierarchical model implemented in `brms` (rather than `rstanarm` since the model is no longer linear). Similarly to the linear model exercise, we found no relationship between the estimated latent spatial abundance coefficients and the estimates of fish abundance provided by the trawl surveys (Fig.\@ref(fig:strct-latent-plot)).


```{r strct-latent-plot, fig.cap="Scaled latent biomass coefficients plotted against paired scaled biomass estimates. Red dashed line indicates a 1:1 fit, solid blue line represents a linear model of the two axes (NOTE FIXING CONVERGENCE ISSUE IN HERE"}
strct_latent_plot
```

### Machine Learning Models

Linear and structural models demonstrate little ability to accurately predicting fish abundance using effort data. We turned to machine learning as a final strategy for predicting fish abundance using fishing effort. To begin with, we tested a range of four different machine learning approaches: random forests [ranger, @Breiman2001], gradient boosted machines (gbm, citation), and multivariate adaptive regression splines (bagged MARS and MARS models, citation). Each of these machine learning models is designed to make use of supplied data to maximize out-of-sample predictive power. However, each model also contains a number of tuning parameters that can only be reliably selected by cross-validation. To that end, we used the `caret` package in R [citation] to perform two repeats of ten fold cross validation across factorial combinations of candidate tuning parameters, and selected the set of tuning parameters that minimized the out-of-sample root mean squared error.

With those tuning parameters in hand, we then utilized the cross-validation routines from those selected tuning parameters to quantitatively compare each of the tuned machine learning models in terms of their out-of-sample predictive power. For each model, we have twenty out-of-sample RMSE estimates. We used the `tidyposterior` package [citation] to fit a Bayesian hierarchical model that, controlling for split effects, estimates the relative effect of each candidate model on out-of-sample RMSE. Based on this analysis, the random forest model (as implemented by the `ranger` package) has the lowest estimated out-of-sample RMSE. As such, we use the random forest, as implemented by the `ranger`, as our candidate machine learning model for the remainder of this paper.

```{r best-ml-plot, fig.cap="Posterior densities of out-of-sample RMSE predicted by `tidyposterior`"}
best_ml_model_plot
```

The other decision to be made within the model fitting process is what resolution of data to use within the fitting process. The finest scale effort data pulled from Global Fishing Watch provides the largest sample size, but also potentially increases the noise in the data. Aggregating the data at coarser spatial aggregations decreases sample size but also may decrease noise. We tested the models at three different spatial resolutions, raw, at 25km^2^ resolution, and 100km^2^ resolution. The 100km^2^ resolution had the highest R^2^ for the training data, and so we will use that as our default resolution for this analysis (Fig.\@ref(fig:ml-res-plot)). Using the 100km^2^ resolution data, the random forest has substantially greater predictive ability within the training data split than any of the linear or structural approaches, with a median R^2^ across the training splits of over 0.5 (Fig.\@ref(fig:ranger-ovp)).

```{r ml-res-plot, fig.cap = "Training data R2 for the random forest (ranger) model at three evaluated spatial resolutions of the data"}
ml_resolution_plot
```

```{r ranger-ovp, fig.cap="Observed vs predicted biomass (A) and R2 values (B) for the fitted random forest models across different evaluated data splits"}
ranger_ovp_plot
```


We can repeat this same resolution procedure to perform one final performance comparison between the random forest and structural models. For each model, we fit the model using the finest resolution data, and then aggregated predictions up to coarser resolutions (Fig.\@ref(fig: strct-v-ml-plot)-A), and refit the model itself using coarser resolution data (Fig.\@ref(fig: strct-v-ml-plot)-B). Using coarser resolution data in the model fitting process improves the predictive power of the structural models somewhat, but the random forest still outperformed the structural approach across all spatial resolutions.

```{r strct-v-ml-plot, fig.cap="Training set R2 from aggregating results of model fit on finest resolution (A) and fitting the model at coarser resolutions (B)"}
strct_v_ml_plot
```

### Value of Information

So far, the model with the greatest predictive power, as measured by the R^2^ of the model within the training data, is a random forest model trained on a random subset of the available data aggregated at a 100km^2^ resolution. Under those conditions, we see training set R^2^ in the vicinity of 0.5. Is this good? Models such as @Costello2012a report R^2^ near 0.4, so 0.5 would appear to be a respectable value. However, the explicit purpose of this analysis is to determine the value of the effort data supplied by Global Fishing Watch for estimating fish abundances. To get at this, we can compare the predictive power of the Global Fishing Watch based model to an alternative model for estimating fish abundance using different sources of information. There are clear theoretical reasons to believe that fishing effort should respond to and affect fish biomass. However, the environment also plays a substantial role in driving fish dynamics, both in abundance and spatio-temporal distribution [@Szuwalski2016; @Munch2018]. A model of fish abundance based solely on environmental drivers makes at least as much conceptual sense as a model based on effort and fleet characteristics then.

Based on this idea, we pulled globally estimates of chlorophyll concentrations (a measure of primary productivity, citation), along with sea surface temperature, and bathymetry, from the National Oceanic and Atmospheric Administration ERDDAP platform (citations). We paired these data with non-effort based data pulled from Global Fishing Watch (distance from shore, MPA status), since these data are not part of the novel effort data provided by Global Fishing Watch. We then refit the machine learning models (since the structural models require effort data) using only the environmental data, and using both the environmental data and effort data (following identical fitting procedures across all runs). This allows us to assess the change in predictive power (as measured by R^2^ of the training data) that including effort data provides.

Comparing effort data alone vs environmental data alone, we see that the relative value of information of the effort data is in fact negative. Meaning, the environment-alone model substantially outperformed, in terms of training data R^2^, the effort-alone model. If the effort data by themselves are not as predictively useful as environmental drivers, are the effort and environmental data together worth more than the sum of their parts? Our results suggest that in fact they are not; combining the effort data with the environmental data provides nearly identical performance to using just the environmental data (Fig.\@ref(fig:voi-plot)).

```{r voi-plot, fig.cap="Differences in R2 or tuned random forest model relative to R2 obtained from only using environmental data (negative implies worse performance than environmental data only)"}
voi_plot
```
### Confrontation with Testing Data

The preceding steps have determined that the best model, in terms of training dat R^2^, is a random forest fitted with data aggregated at the 100km^2^ resolution, using both environmental and effort data. The end goal of this model though is not to predict data within the training set; instead we would hope to use this model to help us understand fish abundance in places outside of the data used to train the model, either in space (i.e. new locations) or time (periods not covered by the trawl surveys). As discussed earlier, the decision of what model to use must be made by examining the training data (and splits of the training data) alone. Now that we have used the training data to select a model that the evidence suggests will have the highest chance of performing well out of sample (remember that even within the training data, the random forest looks to avoid overfitting), we can now confront our selected model with the held out testing data. We also include the structural models in this comparison to see if the structural assumptions of these models provide it an advantage in out-of-sample prediction, though the structural model does not have the same built-in resistance to overfitting to the training data that the machine learning models have.

The predictive performance of our candidate models against the testing data indicate that the decisions based on the training data were well founded. Inclusion of the effort data did not improve the performance of the models on the testing data, and across all cases the machine learning approaches still outperformed the structural models. Looking at just the random test-training splits then, our results would seem to show that a random forest based largely on environmental drivers is a good out-of-sample predictor of fish abundance. Before we start replacing stock assessments with remote sensing of environmental drivers though, we should look at the out-of-sample performance of the models trained on other data splits.

Under the "historic" data split, the training split consists of observations from 2012-2014 and the testing split from 2015-2017. Under this split the performance from the training to the testing split drops off much more dramatically than under the random split, showing that predicting new years is a much more challenging task for the model than filling in gaps within a year. Similarly, we see that a model trained on data off of the Washington/Oregon coast alone is almost completely useless as a predictor of fish biomass.

```{r test-training-plot, fig.cap="R2 for testing and training splits across candidate variables and models"}
test_training_plot
```

## Discussion

The goal of this project was to determine the value of the effort data provided by Global Fishing Watch in estimating fish abundance in space and time. We accomplished this by first determining through a set of fitting routines (models, tuning parameters, resolution) that, given the training data, appeared to provide the highest likelihood of performing well, in terms of predictive ability, both in and out of sample. This process found a random forest tuned on coarser spatial resolution data to be the "best" model. From there, we were able to estimate the value of information of the effort data by comparing the predictive power of the selected model using effort data, as compared to the same selected model using only non-effort based data. This analysis showed that the effort data provides little predictive power beyond that provided by the non-effort data alone.

That the machine learning models outperformed the liner and structural models comes as no surprise: machine learning models are designed explicitly for prediction and should be expected to do well at the task. However, they have two central weaknesses that made the evaluation of alternative predictive models important. The first is that they lack any structural assumptions, and therefore are relatively uninterpretable. Structural models, such as the bio-economic approach taken here, require assumptions about specific functional forms, but as a result provide a means for rationalizing results, e.g. by providing parameter estimates that can be evaluated and interpreted using statistical methods and theoretical knowledge (e.g. the meaning of a cost coefficient can be understood and our confidence the value of that coefficient estimated). All else being equal, we would clearly prefer to have an interpretable model than a black box. To that end, when prediction is the objective, we can estimate the "cost" of that interpretability by comparing the predictive ability of a machine learning approach that maximizes predicability of at the expense of interpretability to a structural model that seeks to maximize the likelihood of the data conditional on its assumptions. While, during the training phase, the structural model is unlikely to outperform the machine learning approach, if it comes close, the price in predictability may be well worth the gain in interpretation. In our case though, the machine learning approach so outperformed the structural approach that they cannot be outweighed by the interpretability of the structural approach. This does not mean that the broad concept of the ideal-free distribution that is at the core of the structural model is inherently incorrect, but that that particular model is not suitable for the aggregation of the data as they stand. It is entirely possible that finer resolution effort and abundance data (e.g. the logbook data utilized in @Miller2016) would produce better performance from the structural model. But, our results show that a structural bio-economic modeling approach is not appropriate for using the effort data supplied by Global Fishing Watch to estimate fish abundance.

Vastly out-of-sample prediction is a second major problem with machine learning models, and the random forest model selected through this process in particular. A properly specified and estimated structural model provides a clear process for predicting outcomes in situations that are far outside of the scale of the training data. If for example the structural model is a simple linear regression with a slope and intercept, trained on one dependent variable on the range one to twenty, and we then confront our fitted model with a new dependent datapoint with value of 1,000, our estimated slope and intercept allows us to easily provide a prediction for this new data point. The random forest is able to do this process as well, but lacks a clear mechanism xx for doing so. A random forest works by fitting a forest of regression trees, each of which, in the case of continuous predictors, break the predictors into a series of bins. Therefore, the predicted outcome for a dependent variable 100 times greater than any value used in the training value will be more or less the same as the prediction for the largest value of the dependent value in the training data (i.e. the new data point will fall into the "greater than some cutoff" bin).  If there is some continuous relationship between the dependent variable and the outcome, this prediction may be severely biased. Because of that, machine learning models such as random forests are best at "filling in the gaps" for data fitting within a defined parameter space, and can struggle when fit on one parameter space and applied to a vastly different parameter space.

We checked for the possibility of this problem by, post selection of the random forest model, examining the predictive ability of that model on both the testing data and the held out training data. Our results show that this flaw of random forests is indeed a problem here: the model does very well at predicting "out of sample" points that are simply random omissions from the complete database. A model trained on the years 2012-2014 and used to predict 2015-2017 (the "future" training-test split) performs worse than the random split, and a model trained on Washington/Oregon data and used to predict California losses roughly half of its predictive power. These results show that there are predictable relationships between fish biomass and  environmental, and to a lesser extent effort, data, but that these relationships do not easily export to new time periods or locations.

What should we make of the relative lack of predictive value of the effort data, as compared to the environmental data? It is critical to note that this is not to say that the effort data alone does not have predictive power, at least within the rough survey region and time period on which the models are trained. R^2^ values for a random forest trained on a random subset of the data to predict fish biomass were near 0.5, both for the training and testing splits. The effort data alone are capable predictors. But, if the question is what additional predictive power do these effort data provide us that we could not have obtained from other data streams, such as environmental data, the answer is not much.

While this is not the result that we might have hoped for, it is perhaps not surprising for two reasons. The first is that perhaps this is simply an indication of the long-understood challenges of using effort data alone to make meaningful inferences on the status of fish stocks: more fishing might mean abundant fishing grounds or overexploited locations where the fishing is cheap. A second and more interesting reason though may lie in the nature of the information used by fishermen to make their decisions. While many bio-economic modeling exercises assume perfect knowledge of the location and amount of fish stocks for simplifying reasons[citations], in reality of course the choice of where and how much to fish is an uncertain and complicated process, based on past experience and shared knowledge. Part of that decision making process is directly related to observing and reacting to the precise type of environmental drivers included in this analysis. Fishermen understand the preferred depth and temperature contours of their target species, and areas of substantial upwelling, often marked my increased chlorophyll concentrations, have long been known to be productive fishing grounds [citations]. So, while the model hypothesized that combinations of environmental and effort data might provide a signal worth more than the sum of their parts, our results suggest in fact with regards to predicting biomass, the effort data a noisier reflection of the environmental data, but further muddied by the myriad other factors affecting fishing decisions, including costs, regulations, experience, and safety.

The effort data provided by Global Fishing Watch present a novel and massive influx of information that shed light on a variety of different factor affecting our oceans, including the footprint of global fishing [@Kroodsma2018] to the estimates of the profitability of different fishing regions [citation]. This project evaluated the extent to which these data could be used to improve fisheries management by helping estimate fish abundances associated with different effort patterns. We found that these effort data can be used for predicting fish biomass, but that a) the manner in which effort is related to abundance, at least at these aggregated resolutions of multiple gears and multiple species, is poorly described by classical bio-economic models, and that b) while machine learning models were able to provide much greater predictive power, the effort data provided little additional predictive value over other globally available environmental datasets. Further work utilizing effort data derived from Global Fishing Watch in stock assessment will need to find ways of more closely matching effort data with their targeted species, or shift attention from using effort as an indicator of biomass towards using it as a prior on the evolution of fishing mortality rates. 
