---
title: "Estimating Fish Abundance from the Behavior of Fishing Fleets"
authors: "Dan Ovando, Chris Costello, Steve Gaines, Ray Hilborn, Jim Thorson"
output:
  bookdown::pdf_document2: null
  bookdown::html_document2: null
  html_document: default
linkcolor: blue
bibliography: dissertation.bib
biblio-style: apalike
---

```{r, include=F}
# knitr::opts_chunk$set(message = F, warning = F, echo = F ,
#                       fig.width = 10, fig.height = 8, eval = T, cache = F)
```


```{r}
library(brms)
library(sf)
library(viridis)
library(scales)
library(rstan)
library(patchwork)
library(tidyverse)
extrafont::loadfonts()


functions <- list.files(here::here("functions"))

walk(functions, ~ here::here("functions", .x) %>% source()) # load local functions

run_name <- 'd1.1'

run_dir <- here::here('results',run_name)

load(file = "/Users/danovando/PhD/skynet/results/d1.1/skynet-plots.Rdata")

```


## Executive Summary

The goal of this study is to determine the ability of Global Fishing Watch (GFW) derived data on fishing activity to predict fish abundance as estimated from fishery independent research surveys.

We do this by testing the ability of structural and machine learning models to use GFW data to make accurate out-of-sample predictions.

Current results suggest that structural approaches are not capable of this task, but the machine learning approaches show potential.

The model can predict absolute and relative spatial abundance with some accuracy. The model can also capture overall temporal trends, but struggles to perform for time periods excluded from the model training.

### Next Steps

  - Fitting the model using only GFW derived data (no environmental covariates)
      - might degrade in-sample performance but improve out-of-sample performance

  - Fitting the model using only unfished species
      - If we are truly capturing a "fishing signal", the the performance of a model fit to unfished species should degrade substantially (though we would still expect it to work somewhat due to covariance between fished and unfished species)

  - Getting our hands on a much more "out of sample" dataset: North Sea, Mediterranean, West Africa?

  - Determining a method for benchmarking performance: How good is good enough?
      - Compare performance of Catch-MSY/surplus production model estimates of abundance trends?
      - ....?

  - Do one more pass at a "keep it simple stupid" approach: Get catch data for the survey regions for all the species in the surveys, take total catch divided by total effort from GFW, see how that index of multi-species abundance compares to survey based estimates.



## Introduction

<!-- Is the paper about global status? At this point, no it's not. It's about linking behavior to abundance, so lead with that. Once it works, then you can start thinking about global applications.  -->

Fishing activity has a massive footprint, spanning the globe and ranging from small-scale subsistence practices up to huge industrial operations. Across these scales though, fisheries share a common underlying incentive structure: fishermen desire some utility derived from capturing fish (e.g. some combination of food, income, and cultural activity) and tune their fishing activities in order to try and maximize that utility, subject to the constraints of the world. As time goes on, these fishing actions affect fish stocks, for example through reductions in abundance or mean length, causing the behavior of the fishermen to be updated. In short then, fishing and fish are part of a dynamic system, in which the behavior of each affects the behavior the other.

These dynamic links between fishing fleets and fish populations are a critical part of fisheries management. @Gordon1954 predicted that in the absence of property rights or access restrictions, these bio-economic dynamics would result in the fishery reaching an open-access equilibrium at which net profits in the fishery are driven to zero, often resulting biological overfishing of the stock. This thinking was central to the bleak predictions of the "tragedy of the commons" [@Hardon1968]. While @Ostrom1990 demonstrated that the evolution of natrual resource systems such as fisheries are driven by a more complete and complex set of drivers than pure profitability, the critical link between the dynamics of fish populations and fishing communities remains in the form of social-ecological systems [@Ostrom2009].

Understanding the dynamics of these social-ecological systems is critical to effective fisheries management. Bio-economic theory and empirical evidence helps us design and implement policies that best achieve societal objectives, by allowing us to model what the potential impacts of policy choices may be. This process has underpinned the recent expansion in the use of rights-based fisheries management [e.g. @Costello2008; @Costello2017; @Deacon2007; @Grainger2016; @Kaffine2008; @Wilen2012; @Cancino2007; @Grafton1996; @Squires2016], and increasingly is used in the management-strategy-evaluation [MSE, @Butterworh2007; @Punt2016a] process. @Nielsen2017 provides a thorough review of models currently utilizing bio-economic modeling to model and guide the policy implementation process.

These works demonstrate a rich history of thinking about how fishing fleets respond to incentives and population dynamics. However, the vast majority of the literature in this field is focused on predicting fishing effort as a function of fish populations; very little research has reversed this question and asked, what do the dynamics of fishing effort suggest about the state of fish populations? If fisheries are indeed a coupled bio-economic system, then just as we believe the dynamics of fishing effort should be predictable from fish abundance, fish abundance should be predictible from fishing effort [@Sugihara2012]. The reason for this omission is likely driven by the realities of fisheries management. Bio-economic modeling is almost exclusively used in the MSE phase of fisheries management, rarely in the stock assessment phase itself. This is partly due simply to the history of stock assessment as a biological science [@Smith1994], and partly due to data realities. Historically, most fisheries with say accurate and complete records of fishing effort would also have records of more directly useful (from the perspective of traditional stock assessment) indicies such as catches and catch-per-unit-effort. In these cases the catch data, combined with catch-per-unit-effort (CPUE) data, can provide a clearer signal of the state of a fish stock (with the catch data providing information on the scale of abundance and CPUE data on trends) than the effort data.

<!-- Many fisheries management policies are based around first achieving a biological objective (e.g. not overfishing the stock), and policies are then evaluated in terms of alternative strategies for achieving that objective.  -->

While effort dynamics have not been historically used to understand fish populations, it is reasonable to believe that given sufficient data, it should be possible to infer something about the state of a fish population as a function of the behavior of fishing fleets. This idea is in many ways analogous to earlier research linking the behavior of sea birds and the location of their prey [citation]. While this idea is grounded in sound theory, the actual form of link between fishing effort and fish abundance is far from clear. High levels of fishing effort could reflect high abundance of fish in the earlier years of a fishing ground, or could reflect an overfished but easily accessible region. We hypothesize that all else equal, fishermen would like to maximize their utility from fishing, but the varied and complex nature of these individual utility functions, combined with the shifting and uncertain nature of the natural world, make the structure of the link between fishing behavior and fish abundance a complicated question.

It is therefore far from obvious what simply observing the effort of fishing tells us about the abundance of fish. This uncertainty is in part why raw effort trends are rarely integrated  into the models we use to estimate the status of fish stocks [citations].
However, recent advances in technology provide reason to rethink what we could learn about fish abundance from fishing behavior. Global Fishing Watch [GFW, presented in @Kroodsma2018], provides a new and vast picture of fishing effort around the planet in near real time. GFW provides high resolution data detailing factors such as the number and horsepower of fishing vessels active in a given region and the distance that fishermen are moving from port to fish. We pair these data with fishery-independent research surveys compiled by the [FishData](https://github.com/James-Thorson/FishData) package in R [@RCoreTeam2018] to ask, can the GFW derived effort data be used to predict the abundance of fish?

We break this assessment into a series of discrete phases.

  - We assess the relationships between total effort and biomass indicies in order to see if, simply put, more fishermen = more or less fish

  - We assess whether total effort can be paired with regional catch data to create an index of catch-per-unit-effort which in turn provides a reasonable index of fish abundance

  - We consider the ability of a series of structural economic models, based on bioeconomic theory, to predict fish abundance as a function of fishing behavior

  - We compare the predictive performance of the structural models to a suite of machine-learning models that utilize the same GFW data, and available covariates if desired, to predict fish abundance

  - Lastly, we consider the predictive value of GFW derived information relative to other globally available indicies (e.g. sea surface temperature and chlorrophyl)


<!-- Answering this question can help us address several critical gaps in our knowledge of marine ecosystems. First, the majority of fisheries in the world lack data for formal stock assessment, limiting our ability to understand and manage these critical resources. To address this gap, numerous publications have attempted to use catch data supplied by the United Nations Food and Agricultural Organization [@cite] to provide some understanding of the status of the world's fisheries, as this bas been the only global database available [@Costello2012; @Costello2016;@Rosenberg2017]. GFW's effort data provides a new opportunity for understanding these fisheries.  Second, while we are increasingly interested in ecosystem-based fisheries management, many of our data-collection practices are still centered around single-species management [@cite]. GFW's effort data provides a potential link to status of all the fish stocks targeted by the fishing gears and fleets covered by GFW. Third, climate change is predicted to change the distribution and range of fish stocks around the planet [@cite]. GFW's data could allow us to track shifts in fishing effort and fish abundance potentially influenced by climate change [@citation]  -->


<!-- While the field has made dramatic advances in our ability to assess fisheries, by and large we have found two solutions to this problem: Fit highly complex integrated statistical models to diverse data streams [cite fancy stock assessment technique], or utilize increasing levels of statistical wizardry to try and squeeze more information out of limited data [what has lately been termed Data Limited Stock Assessments, or DLAs; cite]. The explosion of DLAs has been both promising and concerning. The majority of fisheries in the world lack the resources for fully integrated stock assessments, and so depend on this world of "data limited stock assessment". While there has been tremendous growth in this field, nearly all DLAs rely on the same streams of information that would have been available to a fisheries scientist in the 1800s: lengths [cite], captures [cite], and catch per unit effort[cite]. -->

<!-- It is high time that we expanded not only the statistical methods of fisheries, but also the actual sources of data available for monitoring fish populations. GFW.... etc.  -->


## Results

### Linear Models

Before heading down the statistical rabbit hole, we can simply examine how well linear transformations of effort predict abundance. This has an intuitiatve aspect to it; we can hypothesize that the reason that more fishing occurs in the challenging waters off Alaska than Santa Barbara is that there are higher volumes of valuable fish in that area. However, we could also imagine a scenario where fishing effort is concentrated in overfished but inexpensive grounds, leaving higher fish abundance in more remote areas that are not economical to fish. 

Looking at the effort and abundance indicies, we see some evidence of a "more fishing where there's more fish" hypothesis. Across each of the survey regions, aggregating up to a 200 km^2^ area, there is a positive correlation between the total hours of applicable fishing observed by GFW in that area and the total estimated revenues available in that area (estimated by the sum of the density per knot times the area of that knot times the price of each price of each species in that knot witihn the area encompased by the 200 km^2^ box). However, the relationship is far from clear, with subtantial variation around the mean slope for each region. In addition, we see if all one knew was the total amount of fishing hours, the magnitude of the fishing oppotrunity that those hours might correspond to varies substantially. (Fig.\@ref(fig:effort-v-fish)).

```{r effort-v-fish, fig.cap = "Total fishing hours plotted against fish revenue density. Available revenue is calculated as the sum of the densities of each species times their respective ex-vessel price. Each point represents a 200km^2^ area"}

e_v_ed_plot

```


This coarse data analysis suggests that there may be a relationship between total fishing effort and the value of the fishing opportunity witihn a region, but certainly not a clear enough relationship to serve as a reliable predictor of fishable biomass. That effort levels alone are not clearly informative is not very surprising. What though do we learn by pairing effort data with locally available catch data to construct a CPUE index? CPUE can, under the appropriate circumstances, serve as an index of relative abundance. To create a GFW derived CPUE index, we pulled catch data for the relevant regions and species from three different databases: the RAM Legacy Stock Assessment Databse [@Ricard2011], the NMFS commercial landings database [citation], and the FAO's capture production database [citation]. Pairing these catch data with the the GFW effort data gives us a timeseries of aggregate CPUE for a given region. We then compared these CPUE trends to trends in biomass provided by the RAM database and from the processed trawl survey used throughout this study. We see no clear agreement between these two metrics, and if we take the RAM and survey derived indicies of abundance as true, the GFW derived CPUE index is in some cases provides an exactly oppositive index of abundance from the true value (Fig.\@ref(fig:gfw-cpue))

```{r gfw-cpue, fig.cap="GFW derived CPUE (yellow lines) and assessments of abundance (blue) for the US West Coast and Eastern Bering Sea regions"}

wc_cpue_plot + ebsbts_cpue_plot

```


The previous models are based on the idea of using effort to predict fish abundance. We could also however flip this equation, and ask what fish abundance best explains the observed effort patterns? We test this by estimating a linear model of the form

$$effort_{i,y} \sim cost_{i,y} + space_{i}+\epsilon_{i}$$

$$space_i\sim normal(region_{i},\sigma_{space})$$

Where for an observation at location *i* and year *y*,  *cost* is made up of variables such as distance from port, distance from shore, depth, and mean length of vessels at that observation, and *space* is a hierarchichal parameter representing a random effect for that location (under the assumption that correlations exist in site effects within a given survey region). Our hypothesis here is that by controlling for factors that should explain a substantial portion of the cost drivers for fishing at a given location (distance from shore/port, depth, and the size the vessels fishing there), the effect of the fish abundance expected by the fishermen at that location will be captured by the *space* variable. 

We used a Bayesian hierarchichal model implemented through the `rstanarm` package to estimate this linear model. We then extracted the estimated *space* coefficients, and comparted them to the fish biomass estimated at that location by the relevant fishery independent survey. We find no correspondence between the estimated space effects and the fish biomasses estimated from the trawl surveys. This does not mean that such an approach might not work given sufficient data, but with the avaialable covariates either too many other important factors besides biomass are being absorved into the *space* coefficients, or the survey biomass is a small component of the decision making process for a given fishing location (Fig.\@ref(fig:linear-latent-plot)).  

```{r linear-latent-plot, fig.cap="Scaled space coefficients plotted against paired scaled biomass estimates. Red dashed line indicates a 1:1 fit, solid blue line represents a linear model of the two axes"}

linear_latent_plot

```


### Structural Models 

Since raw effort and effort derived CPUE indicies do not appear to be valid methods for estimating abundance, we now turn to more detailed modeling approaches to utilize effort data to predict abundance. Our structural modeling approach follows a standard bio-economic framework, as layed out in @Miller2016. @Miller2016 used a structural modeling approach in part to estimate the quota prices in the US West Coast grouddfish trawl fishery individual fishing quota (IFQ) program, using data on logbook reported CPUE, prices, amd variable and fixed costs (labor, fuel, etc.). Making the assumption that for an appropriate fleet unit and time period marginal profits are equal in space, @Miller2016 then estimates quota prices for different species that, given their other data, rationalize the observed distribution of effort in the fishery. 

We follow a similar approach but rearrange the problem to estimate abundance instead of effort. We do so by fitting a model that estimates fishing costs as a function of distance from port, distance from shore, vessel size, and hours fished. Fishing costs are estiamted by finding the costs that, given the observed distribution of effort, rationalize the biomass reported by the fishery independent trawl surveys....see methods...

Fitted cost coefficients 

The biomass-predicting structural model shows limited predictive ability within the training set. For all but three of the training datasets, the R^2^ was less than 0.1... 

What else to say?


```{r cost-coefs}

```

```{r strct_o_v_p, fig.cap= "Observed vs predicted biomass (A) and R2 values (B) for the fitted structural models across different evaluated data splits"}

strct_ovp_plot

```

We can also invert this idea in the same manner as we did with the linear model, and rather than estimating cost coefficients that explain the observed fish abundance, given observed efforts, estimate cost coefficents and latent spatial parameters representing abundance that explain the observed effort. We estimated this model using a Bayesian hierarchichal model implemented in `brms` (rather than `rstanarm` since the model is no longer linear). Similarly to the linear model exercise, we found no relationship between the estiamted latenta spatial abundance coefficients and the estiamtes of fish abundance provided by the trawl surveys (Fig.\@ref(fig:strct-latent-plot)).


```{r strct-latent-plot, fig.cap="Scaled space coefficients plotted against paired scaled biomass estimates. Red dashed line indicates a 1:1 fit, solid blue line represents a linear model of the two axes"}
strct_latent_plot
```

### Machine Learning Models

Linear and structural models demonstrate little ability to accurately predicting fish abundance using effort data. We turned to machine learning as a final strategy for predicting fish abundance using fishing effort. To begin with, we tested a range of four different machine learning approaches: random forests [ranger, citation], gradient boosted machines (gbm, citation), and multivariate adaptive regression splines (bagged MARS and MARS models, citation). Each of these machine learning models is designed to make use of supplied data to maximize out-of-sample predictive power. However, each model also contains a number of tuning parameters that can only be reliably selected by cross-validation. To that end, we used the `caret` package in R [citation] to perform two repeats of ten fold cross validation across factorial combinations of candidate tuning parameters, and selected the set of tuning parameters that minimized the out-of-sample root mean squared error. 

With those tuning parameters in hand, we then utilized the cross-validation routines from those selected tuning parameters to quantitatively compare each of the tuned machine learning models in terms of their out-of-sample predictive power. For each model, we have twenty out-of-sample RMSE estiamtes. We used the `tidyposterior` package [citation] to fit a Bayesian hierarchichal model that, controlling for split effects, estimates the relative effect of each candidate model on out-of-sample RMSE. Based on this analysis, the random forest model (as implemented by the `ranger` package) has the lowest estiamted out-of-sample RMSE. As such, we use the random forest, as implemented by the `ranger`, as our candidate machine learning model for the remainder of this paper. 

```{r best-ml-plot, fig.cap="Posterior densities of out-of-sample RMSE predicted by `tidyposterior`"}
best_ml_model_plot
```

The other decision to be made within the model fitting process is what resolution of data to use within the fitting process. The finest scale effort data pulled from Global Fishing Watch provides the largest sample size, but also potentially increases the noise in the data. Aggregating the data at coarser spatial aggregations decreases sample size but also may decrease noise. We tested the models at three different spatial resolutions, raw, at 25km^2^ resolution, and 100km^2^ resolution. The 100km^2^ resolution had the highest R^2^ for the training data, and so we will use that as our default resolution for this analysis (Fig.\@ref(fig:ml-res-plot)). Using the 100km^2^ resolution data, the random forest has substantially greater predictive ability within the training data split than any of the linear or structural approaches, with a median R^2^ across the training splits of over 0.5 (Fig.\@ref(fig:ranger-ovp)).  

```{r ml-res-plot, fig.cap = "Training data R2 for the random forest (ranger) model at three evaluated spatial resolutions of the data"}
ml_resolution_plot
```

```{r ranger-ovp, fig.cap="Observed vs predicted biomass (A) and R2 values (B) for the fitted random forest models across different evaluated data splits"}
ranger_ovp_plot
```


We can repeat this same resolution procedure to perform one final performance comparison between the random forest and structural models. For each model, we fit the model using the finest resolution data, and then aggregated predictions up to coarser resolutions (Fig.\@ref(fig: strct-v-ml-plot)-A), and refit the model itself using coarser resolution data (Fig.\@ref(fig: strct-v-ml-plot)-B). Using coarser resolution data in the model fitting process improves the predictive power of the structural models somewhat, but the random forest still outperformed the structural approach across all spatial resolutions. 

```{r strct-v-ml-plot, fig.cap="Training set R2 from aggregating results of model fit on finest resolution (A) and fitting the model at coarser resolutions (B)"}
strct_v_ml_plot
```

### Value of Information

So far, the model with the greatest predictive power, as measured by the R^2^ of the model within the training data, is a random forest model trained on a random subset of the available data aggregated at a 100km^2^ resolution. Under those conditions, we see training set R^2^ in the vicinity of 0.5. Is this good? Models such as @Costello2012 report R^2^ near 0.4, so 0.5 would appear to be a respectable value. However, the explicit purpose of this analysis is to determine the value of the effort data supplied by Global Fishing Watch for estimating fish abundances. To get at this, we can compare the predictive power of the Global Fishing Watch based model to an alternative model for estimating fish abundance using different sources of information. There are clear theoretical reasons to believe that fishing effort should respond to and affect fish biomass. However, the environment also plays a substantial role in driving fish dynamics, both in abundance and spatio-temporal distribution [@Szuwalski2016; citation]. An model of fish abundance based solely on environmental drivers makes at least as much conceptual sense as a model based on effort and fleet charactersitcs then. 

Based on this idea, we pulled globally estimates of Chloropphyl A conentrations (an measure of primary productivity, citation), along with sea surface temperature, and bathymetry, from the National Oceanic and Atmospheric Administration ERDDAP platform (citations). We paired these data with non-effort based data pulled from Global Fishing Watch (distance from shore, MPA status), since these data are not part of the novel effort data provided by Global Fishing Watch. We then refit the machine learning models (since the strucural models require effort data) using only the environmental data, and using both the envrionmental data and effort data (following identical fitting procedures across all runs). This allows us to assess the change in predictive power (as measured by R^2^ of the training data) that including effort data provides. 

Comparing effort data alone vs environmental data alone, we see that the relative value of information of the effort data is in fact negative. Meaning, the environment-alone model substantially outperformed, in terms of training data R^2^, the effort-alone XX model. Are the effort and environmental data worth more than the sum of their parts? Our results suggest that in fact they are now; combining the effort data with the environmental data provides nearly identical performance to using just the environmental data (Fig.\@ref(fig:voi-plot)). 

```{r voi-plot, fig.cap="Differences in R2 or tuned random forest model relative to R2 obtained from only using environmental data (negative implies worse performance than environmental data only)"}
voi_plot
```


### Confrontation with Testing Data

All of the preceding steps have determined the best model, in terms of training dat R^2^, is a random forest fitted with data aggregated at the 100km^2^ resolution, using both environmental and effort data. The end goal of this model though would not be to solely predict data within the training set; instead we would hope to use this model to help us understand places outside of the data used to train the model, either in space (i.e. new locations) or time (periods not covered by the trawl surveys). As discussed earlier, the decision of what model to use must be made by examing the training data (and splits of the training data) alone. Now that we have used the training data to select a model that the evidence suggests will have the highest chance of performing well out of sample (remember that even within the training data, the random forest splits the training data in smaller subsets of testing and training data, and tunes its parameters to maximize out-of-sample prediction), we can now confront our selected model with the held out testing data, to see how well it performs. We also include the structural models in this comparison though the structural model does not have the same built-in resistence to overfitting to the training data that the machine learning models have. Our results suggest that our decisions based on the training data were well founded: The model and variables that performed best (in terms of R^2^) also performed best on the completely held out testing data, indicating that it is unlikely that we simply overfit the model during the testing phase (Fig.\@ref(fig:test-training-plot)). 

```{r test-training-plot, fig.cap="R2 for testing and training splits across candidate variables and models"}
test_training_plot
```

## Discussion

The goal of this project was to determine the value of the effort data provided by Global Fishing Watch in estimating fish abundance in space and time. We accomplished this by first determing a set fitting routines (models, tuning parameters, resolution) that, given the training data, appeared to provide the highest likelihood of performing well, in terms of predictive ability, both in and out of sample. This process found a random forest tuned on coarser spatial resolution data to be the "best" model. From there, we were able to estiamte the value of information of the effort data by comparing the predictive power of the selected model using effort data, as compared to the same selected model using only non-effort based data. This analysis showed that the effort data provides little predictive power beyond that provided by the non-effort data alone. 

That the machine learning models outperformed the liner and structural models comes as no surprise: machine learning models are designed explicitly for prediction and should be expected to do well at the task. However, they have two central weakenesses that made the evaluation of alternative predictive models important. The first is that they lack any structural assumptions, and therefore are relatively uninterpretable. Structural models, such as the bio-economci approach taken here, require assumptions about specific funcional forms, but as a result provide a means for rationalizing results, e.g. by providing parameter estimates that can be evaluated and interpreted using statistical methods and theoretical knowledge (e.g. the meaning of a cost coefficient can be understood and our confidence the value of that coefficient estimated). All else being equal, we would clearly prefer to have an interpretable model than a black box. To that end, when prediction is the obejctive, we can esimate the "cost" of that interpretability by comparing the predictive ability of a machine learning approach that maximizes predicability of at the expense of interpretability to a structural model that seeks to maximize the likelihood of the data conditional on its assumptions. While, during the training phase, the structural model is unlikely to outperform the machine learning approach, if it comes close, the price in predictability may be well worth the gain in interpretation. In our case though, the machine learning approach so outperformed the structural approach that they cannot be outweighed by the interpretability of the structural approach. This does not mean that the broad concept of the ideal-free distribution that is at the core of the structural model is inherently incorrect, but that that particular model is not suitable for the aggregation of the data as they stand. It is entirely possible that finer resolution effort and abundance data (e.g. the logbook data utilized in @Miller2016) would produce better performance from the structural model. But, our results show that a structural bio-economic modeling approach is not appropriate for using the effort data supplised by Global Fishing Watch to estimate fish abundance. 

Vastly out-of-sample prediction is a second major problem with machine learning models, and the random forest model selected through this process in particular. A properly specificed and esimtated structural model provides a clear process for predicting outcomes in situations that are far outside of the scale of the training data. If for example the structural model is a simple linear regression with a slope and intercept, trained on one dependent variable on the range one to twenty, and we then confront our fitted model with a new dependent datapoint with value of 1,000, our estiamted slope and intercept allows us to easliy provide a prediction for this new data point. The random forest is able to do this process as well, but lacks a clear mechanism xx for doing so. A random forest works by fitting a forest of regression trees, each of which, in the case of contiuous predictors, break the predictors into a series of bins. Therefore, the predicted outcome for a dependent variable 100 times greater than any value used in the training value will be more or less the same as the prediction for the largest value of the dependent value in the training data (i.e. the new data point will fall into the "greater than some cutoff" bin).  If there is some continuous relationship between the dependent variable and the outcome, this prediction may be severely biased. Because of that, machine learning models such as random forests are best at "filling in the gaps" for data fitting within a defined parameter space, and can struggle when fit on one parameter space and applied to a vastly different parameter space. 

We checked for the possibility of this problem by, post selection of the random forest model, examining the predictive abiility of that model on both the testing data and the held out training data (XX need to include one really out of sample testing training split, e.g. WA-OR)....


Evidence that fishermen may be responding more to the environmental signal than knowledge of the biomass on any given day... especially given timing of the surveys? look into that. 

So what?


## Methods

This study compares the ability of various candidate models to utilize GFW effort data, potentially along with other globally-available covariates, to predict fishery-independent estimates of fish abundance. We judge performance by out-of-sample predictive capability of each of our candidate models, across a range of training-test splits of the data (historic-future, Alaska to West Coast, etc).

### Data

GFW provides our data on the amount and location of fishing effort, along with available observed or estiamted covariates such as vessel size, distance from shore/port, and engine power. Estimates of fish abundance in space and time were obtained from their relevant surveys through the `FishData` package. The following two data sections provide summaries of the data as well as descriptions of relevant data processing steps taken.

#### Survey Data

`FishData` provides access to numerous fishery independent research surveys throughout the world. We use the bottom trawl surveys conducted along the west coast of North America by the National Marine Fisheries Service  (\@ref(fig:knot-map))

  - [Eastern Bering Sea Bottom Trawl Survey](https://www.afsc.noaa.gov/RACE/groundfish/bottom%20trawl%20surveys.php) (ebsbts)
  - [Aleutian Islands Bottom Trawl Survey](https://www.afsc.noaa.gov/RACE/groundfish/bottom%20trawl%20surveys.php) (aibts)
  - [Gulf of Alaska Bottom Trawl Survey](https://www.afsc.noaa.gov/RACE/groundfish/bottom%20trawl%20surveys.php) (goabts)
  - [West Coast Groundfish Bottom Trawl Survey](https://www.nwfsc.noaa.gov/research/divisions/fram/groundfish/bottom_trawl.cfm) (wcgbts)
  <!-- - West Coast Groundfish Hook and Line Survey -->

```{r knot-map, fig.cap='Spatial coverage of fishery independent research surveys (will convert to polygons)'}

knot_plot

```


Each of the surveys contains data on a wide variety of different species, including highly abundanct fished species such as arrowtooth flounder and Alaska pollock, as well as unfished species such as sea anemones. The selected surveys utilize bottom trawl gear, and as such primarily contain bottom-associated species (\@ref(fig:fish-plot)). Surveys are conducted in summer months (July-August for the Alaska surveys and May-October for the West Coast Bottom Trawl Survey).

```{r fish-plot, fig.cap = "Number of positive encounters for the top-10 most observed species in each research survey (XXswitch to common names"}

fish_plot

```

Survey data are provided by `FishData` in their "raw" form (biomass of species per unit of survey effort at a given sampling event). The data therefore require standardization to account for differences in vessel characteristics, spatio-temporal correlation structures, and the presence of zeros in the haul data. This standardization was performed using the [`VAST`](https://github.com/James-Thorson/VAST) package [@Thorson2016a], which implements a spatial delta-generalized linear mixed model across multiple species. In essence, this model leverages the spatio-temporal correlations within and among species (along with to provided covariates) to provide a standardized index of abundance for each of the species in the supplied data.

The result of VAST is a network of "knots" that define polygons of equal density for each species, where the density of each species is measured in units of tons/km^2^ XX (Fig.\@ref(fig:fish-map)).

```{r fish-map, fig.cap = "VAST estimates of density (tons/km^2^) of species observed in bottom trawl surveys"}

  fish_map_plot

```


Surveyed species also vary substantially in their abundance and economic importance. We mark species encountered in the surveys as "fished" if their name, or a synonym for their name identified through the `taxize` package, was found within the global catch records of the Food and Agriculture Organization of the United Nations (FAO citation). For each species, we also obtained price estimates using @Melnychuk2016 (Fig.\@ref(fig:price-plot)). Together, these data provide estimated density for fished species encountered by the US west coast bottom trawl survey program over space and time, along with the associated value of these species.

```{r price-plot, fig.cap = "Mean ex-vessel price (USD/ton) of species in database from @Melnychuk2016"}

price_plot

```

#### GFW Data

GFW data were obtained using the `bigrquery` package in R from the GFW servers (xx check on description here to note back-end access). Data were aggregated to the resolution of year and nearest 0.1 degree latitude and longitude, and for each vessel at this a given location we calculated the total fishing hours spent, average distance from shore, average distance from port, and whether that location is inside an MPA (and if so what kind). We also collected relevant data for that vessel such as its engine power, length, tonnage, and vessel type (trawler, purse-seine, fixed-gear, etc). Together, these data provide fishing effort-related data covering the regions surveyed in our fishery-independent data (\@ref(fig:gfw-map)).

```{r gfw-map, fig.cap="Map of GFW effort data"}
gfw_plot
```

#### Environmental Covariates

We augment with the GFW and FishData data with globally accessible environmental covariates, including remote sensing estimates of

  - Chlorophyll
  - Sea surface temperature
  - Bathymetry

All data were obtained from NOAA ERDDAP portal (https://coastwatch.pfeg.noaa.gov/erddap/index.html), and aggregated as needed to match the resolution of the GFW data (annual and 0.1 degree lat/long resolution). Other environmental data were explored (e.g. wave and wind), but did not have sufficient near-shore coverage for inclusion in the model.

#### Merging GFW and Abudnance Data

The GFW data were merged to the survey data by
  - clipping the GFW to only include observations that occurred within the spatial perimeter of one of the surveys
  - snapping each GFW observation to the nearest (in terms of latitude and longitude distance) knot identified by VAST from the survey data. Since the survey data are generally at a courser resolution than the GFW data, this means that multiple GFW observations are usually associated with any one knot at any one time.

  - Since all surveys are geared towards bottom dwelling species, only bottom-associated gears are included from GFW. In this case that means vessels identified by GFW as trawlers, pots and traps, and set gears (longline and gillnet).

  - We only include species that were observed 10 or more times during each year of the survey to improve model convergence.


The preceding steps were conducted for all constructed databases. We also constructed a series of subset data from this master set, including factorial combinations of

  - Only utilizing GFW from the months in which the relevant NMFS surveys were conducted

  - Only including/excluding commercially exploited species, to test the models performance when fitted to unfished species

  <!-- - We Only including GFW data from the months corresponding to the appropriate survey months -->

  <!-- - VAST provides estimates of abundance in years without surveys by predicting based on spatio-temporal patterns. In order to reduce potential bias of models estimating the VAST process instead of underlying abundance, for now we only include survey data from years in which a survey was actively conducted.  -->




### Candidate Models

We evaluated three classes of candidate models for linking GFW effort to fish abundance:

  1. Linear models
    - These simply link abundance to effort through linear models
  2. Structural economic models
    - These assume that marginal profits are equal across a given spatio-temporal resolution,
    and estimate fish abundances that rationalize the observed fishing Effort
  3. Machine learning models
    - These models make no structural assumptions, but rather
    find the combination of predictor variables that maximize the out-of-sample
    predictive power of the model


We fit a series of candidate models to the data, defined in terms of

- Data Splitting: the data used for training and testing.
    - Random splits (train on a random subset, predict the remaining subset)
    - Spatial splits (e.g. train on Alaska, predict US west coast)
    - Temporal splits (e.g. train on 2010-2013, predict 2014-2015)

- Predictive goal:
    - Absolute abundance in space in time
    - Changes in abundance in space and time

- Model Type:
    - Machine learning (random forests, gradient boosted models)
    - Linear regression (fish abundance ~ fishing effort)
    - Structural modeling (fitting of structural economic models that conform to ideal free distribution)


We fit factorial combinations of each of these candidate models, allowing us to evaluate model performance from a variety of relevant angles. The comparison between the machine learning and structural modeling approaches are worth discussing further for a brief moment. While different in their mechanics, all the candidate machine learning models are black-box models whose sole objective is to maximize the predictive power of the model. The user specifies some model options, but the model decides which data are important and how those data relate to each other. This allows these algorithms to fit highly non-linear models (if the data demand it), without the need to specify an exact statistical or structural form for how variables such as costs, safety, and fish abundance interact to affect fishing behavior. As a result, machine learning models can serve as an effective benchmark for the best possible ability of GFW data to predict fish abundance. The disadvantage is that, while new techniques are emerging for interpreting machine learning model fits, they are inherently black boxes and as such do not permit us to really interpret the meaning of specific coefficients. The lack of a structural theory behind the model may also hamper the ability of these models to predict radically out of sample data (e.g. a machine learning model trained in Alaska may perform terribly in Africa).

This stands in contrast to the structural modeling approach, in which we write out a functional form for the relationship between fish abundance and fishing pressure, and the fit that specified model to the data. This has the advantage of interpretability, but opens us up to errors in model specification. By running both, we can compare the machine learning and structural approach and see how much the interpretability of the structural model "costs" us in terms of predictive power, relative to the benchmark of the machine learning model.

For each model run, the data are split into test and training sets. Within the training sets, the data are further split through v-fold cross validation (using five repeats of ten splits) into assessment and analysis splits, which are used to fine tune model characteristics prior to fitting the model on the full training data and testing performance on the testing data.

Each model was designed using a separate subset of data, so as to isolate  initial model design from the data that will eventually be used to judge model performance.

Below I provide high-level summaries of each of the models, purely to aid in understanding (more detailed descriptions will be provided in future drafts)

### Machine Learning

Two machine learning algorithms, random forests (implemented through the `ranger` package in R) and generalized boosted regression modeling (gbm), were evaluated. Tuning parameters of each method were selected using the `caret` package on an isolated set of data.

A random forest works by fitting a series of regression trees. Each regression tree takes a sub-sample of the training data, and a sub sample of the independent variables provided for model fitting. The algorithm then determines the variable and variable split (e.g. vessel size and vessel size above 30ft) that provides the greatest explanatory power in estimating density of the "out of bag" samples (the part of the training data that were not included in the tree), and creates that as the first node. The next two nodes are selected in the same process, and so on and so forth, down to a specified tree depth tuned through the `caret` package. Each tree provides a high-variance, low bias estimator of densities. The random forest then averages over hundreds of trees to reduce this variance and provide an improved estimate of density as a function of provided covariates. The advantage of this approach is that it makes no assumptions about error distributions or linearity of parameters, and actively pushes back against over fitting by prioritizing out-of-sample prediction, and sub sampling of the provided independent variables (lots and lots of literature on this).

XX Will add in explanation of GBM, but really just a modification of a random forest that helps the model target and improve the fit of parts of the data that the model struggles with

1 Select tree depth, D, and number of iterations, K 2 Compute the average response, y, and use this as the initial predicted value for each sample 3 for k = 1 to K do 4 Compute the residual, the difference between the observed value and the current predicted value, for each sample 5 Fit a regression tree of depth, D, using the residuals as the response 6 Predict each sample using the regression tree fit in the previous step 7 Update the predicted value of each sample by adding the previous iteration’s predicted value to the predicted value generated

(Page 205, Kuhn & Johnson).


### Structural Model

We fit a structural model in the manner of @Miller2016 to the data. The key of this model assumption here is that fishermen conform to an ideal free distribution, and so marginal profits are equal across appropriate groups in space and time, at the unit of year and "knot", a knot being a geographic region of constant abundance estimated per the methods of @Thorson2016a.

Following @Miller2016, we consider marginal profits per unit effort as being

$$\pi_{y,k} = pCPUE_{y,k}e^{-qE_{y,k}} - c_{y,k}$$

where for year *y* at knot *k*, *p* is price, CPUE is the index of abundance prior to any fishing effort occurring (our index of abundance, *q* is catchability, *E* is effort, and *c* are variable costs).

@Miller2016 was primarily interested in estimating quota price aspects of *c*, taking as data *p*, *CPUE*, *E*, and other components of *c* (fuel, labor, ice, etc.). We are instead interested in estimating CPUE as a function of other variable, and so we can rearrange this equation as

$$ CPUE_{y,k} = \frac{\pi_{y,k} + c_{y,k}}{pe^{-qE_{y,k}}}$$

Similar to @Miller2016 we assume for now that $\pi_{y,k}$ is zero, though this is clearly not accurate given that many of the fisheries sampled by this model are highly regulated and in some cases rationalized (however, changing $\pi_{y,k}$ to positive values had little effect on the fit of the model). *p* is taken from @Melnychuk2016, and $E_{y,k}$ is observed from GFW.

That leaves *q* and *c* as unknown parameters. While we do not have the high resolution logbook data available to @Miller2016, we could certainly obtain data on fuel and labor prices for this model. However, at this time, I simply assume that $c_{y,k}$ is a function of the distance of a knot *k* from port, the mean vessel size used at knot *k* in year *y*, and the interaction of these terms (under the assumption that cost of large vessels increase with distance).

$$ c_{y,k} = \beta_{1}pd_{k} + \beta_{2}vs_{y,k} + \beta_{3}pd_{k}vs_{y,k} $$

Where $pd_{k}$ is the distance from port of knot *k* and $vs_{y,k}$ is the mean vessel length observed fishing at knot *k* in year *y*.

We fit the model through maximum likelihood using Template Model Builder (TMB).

<!-- $$[q, \beta_{1}, \beta_{2},\beta_{3} | log(CPUE_{y,k}()] \propto normal(log(CPUE_{y,k}) | f(q, \beta_{1}, \beta_{2},\beta_{3} ),\sigma)$$ -->

<!-- Repeated initial values and a jittered while loop were used to ensure model convergence (though a full Bayesian implementation will improve on this at a later date). We could of course again include either informative priors on the $\beta$ parameters based on fuel and labor costs, or directly incorporate those data, but this provides a useful starting point.  -->

```{r, eval = F}

p <- 10
cpue <- 100
q <- .1
effort <- 0:100
cost <- 200

mp <- p * cpue * exp(-q*effort) - cost

plot(effort,mp)

```

### Linear Models

As a counterbalance to the relative complexity of the machine learning and structural approaches, we also explore a simple linear model of the form

$$log(density_{i}) = \beta_0 + log(\beta_1effort_i)$$

We include this mostly to explore the basic evidence for a hypothesis that more effort means high fish density.


- Fishermen responding to environmental cues?
- What does this say about spatial bioeconomic modeling?
- Relative importance of environment?




## Summary

So far the GFW data show a limited but promising ability to predict CPUEs outside of sample (both through K-fold cross validation and through omission of the most recent data).

There is clearly a lot of work to be done from here, and I think that improvements in model fit can be achieved through re-scaling of coefficients (e.g. we don't really care about absolute CPUE, but rather the trend). We will also need to explore the ability of the model to perform across different aggregations of fleets and species. The structural model can also be improved by integration of additional data on fuel and labor costs.

Early results suggest then that this idea isn't insane. The most rigorous approach then will be a careful development and comparison of machine learning and structural models. However so far the machine learning approach is in the front of the pack by a decent distance.


## Literature notes

see @vanPutten2012 for references supporting fish abundance as a driver of economic activity

@Gillis2012 for an alternative economic model to the ideal free distribution. Also has great citations for behavior tracking abundance for snow crabs, as well as failures for other fisheries

