---
title: "Vastify Trawl Survey Data"
output: html_notebook
---

```{r}
knitr::opts_chunk$set(echo = F, message = F, warning = F)
library(tidyverse)
library(VAST)
library(ThorsonUtilities)
library(demons)
library(sf)
library(ggmap)
library(FishData)
library(viridis)
library(compiler)
library(TMB)

```
Goal here is to use the VAST package to "interpolate" between the observed survey locations (I know interpolate isn't the word, but if not, then what the hell is it)


```{r}

ebs_trawl <-
  FishData::download_catch_rates(survey = "EBSBTS", species_set = 50) %>%
  set_names(tolower(colnames(.)))

  


```

```{r wrap-hook, echo=FALSE}
# FROM: https://github.com/yihui/knitr-examples/blob/master/077-wrap-output.Rmd
library(knitr)
hook_output = knit_hooks$get('output')
knit_hooks$set(output = function(x, options) {
  # this hook is used only when the linewidth option is not NULL
  if (!is.null(n <- options$linewidth)) {
    x = knitr:::split_lines(x)
    # any lines wider than n should be wrapped
    if (any(nchar(x) > n)) x = strwrap(x, width = n)
    x = paste(x, collapse = '\n')
  }
  hook_output(x, options)
})
# TRIGGERED USING `linewidth=60`
```



Prepare VAST

```{r}
Version = "VAST_v2_4_0"
Method = c("Grid", "Mesh")[2]
grid_size_km = 25
n_x = c(100, 250, 500, 1000, 2000)[1] # Number of stations
Kmeans_Config = list( "randomseed"=1, "nstart"=100, "iter.max"=1e3 )  


FieldConfig = c("Omega1"=1, "Epsilon1"=1, "Omega2"=1, "Epsilon2"=1) 
RhoConfig = c("Beta1"=0, "Beta2"=0, "Epsilon1"=0, "Epsilon2"=0) 
OverdispersionConfig = c("Vessel"=0, "VesselYear"=0)
ObsModel = c(2,0)  

Options =  c("SD_site_density"=0, "SD_site_logdensity"=0, "Calculate_Range"=1, "Calculate_evenness"=0, "Calculate_effective_area"=1, "Calculate_Cov_SE"=0, 'Calculate_Synchrony'=0, 'Calculate_Coherence'=0)

Region <- 'Eastern_Bering_Sea'


DateFile = file.path(getwd(),'VAST_output/')
  dir.create(DateFile)
  
  
  

alaska_pollock <- ebs_trawl %>%
  filter(sci == sci[1]) %>% 
  rename(Year = year, Lat = lat, Lon = long, Catch_KG = wt) %>% 
  mutate(AreaSwept_km2 = 1, Vessel = 'missing') %>% 
  select(Year,Lat,Lon,Vessel, AreaSwept_km2, Catch_KG)

qmplot(x = Lon, y = Lat, data = alaska_pollock, color = log10(Catch_KG)) +
  scale_color_viridis() +
  facet_wrap(~Year)

  Record = ThorsonUtilities::bundlelist( c("alaska_pollock","Version","Method","grid_size_km","n_x","FieldConfig","RhoConfig","OverdispersionConfig","ObsModel","Kmeans_Config") )
save( Record, file=file.path(DateFile,"Record.RData"))
capture.output( Record, file=paste0(DateFile,"Record.txt"))


```

Prepare for extrapolation

```{r}

  strata.limits <- data.frame('STRATA'="All_areas")

  Extrapolation_List = SpatialDeltaGLMM::Prepare_Extrapolation_Data_Fn( Region=Region, strata.limits=strata.limits )
  
  
  #qmplot(x = Lon, y = Lat, data = Extrapolation_List$Data_Extrap)
```



```{r}
alaska_pollock <- alaska_pollock %>% 
  na.omit() %>% 
  mutate(Vessel = as.factor(Vessel)) 

Spatial_List = SpatialDeltaGLMM::Spatial_Information_Fn( grid_size_km=grid_size_km, n_x=n_x, Method=Method, Lon=alaska_pollock[,'Lon'], Lat=alaska_pollock[,'Lat'], Extrapolation_List=Extrapolation_List, randomseed=Kmeans_Config[["randomseed"]], nstart=Kmeans_Config[["nstart"]], iter.max=Kmeans_Config[["iter.max"]], DirPath=DateFile, Save_Results=FALSE )
# Add knots to Data_Geostat
alaska_pollock = cbind( alaska_pollock, "knot_i"=Spatial_List$knot_i )

```



```{r}
TmbData = Data_Fn("Version"=Version, "FieldConfig"=FieldConfig, "OverdispersionConfig"=OverdispersionConfig, "RhoConfig"=RhoConfig, "ObsModel"=ObsModel, "c_i"=rep(0,nrow(alaska_pollock)), "b_i"=alaska_pollock[,'Catch_KG'], "a_i"=alaska_pollock[,'AreaSwept_km2'], "v_i"=as.numeric(alaska_pollock[,'Vessel'])-1, "s_i"=alaska_pollock[,'knot_i']-1, "t_iz"=alaska_pollock[,'Year'], "a_xl"=Spatial_List$a_xl, "MeshList"=Spatial_List$MeshList, "GridList"=Spatial_List$GridList, "Method"=Spatial_List$Method, "Options"=Options )

TmbList = Build_TMB_Fn("TmbData"=TmbData, "RunDir"=DateFile, "Version"=Version, "RhoConfig"=RhoConfig, "loc_x"=Spatial_List$loc_x, "Method"=Method)
Obj = TmbList[["Obj"]]

```

```{r}
Opt = TMBhelper::Optimize( obj=Obj, lower=TmbList[["Lower"]], upper=TmbList[["Upper"]], getsd=TRUE, savedir=DateFile, bias.correct=FALSE )
```


Finally, we bundle and save output
```{r save_results, linewidth=60}
Report = Obj$report()
Save = list("Opt"=Opt, "Report"=Report, "ParHat"=Obj$env$parList(Opt$par), "TmbData"=TmbData)
save(Save, file=paste0(DateFile,"Save.RData"))
```


```{r}
SpatialDeltaGLMM::Plot_data_and_knots(Extrapolation_List=Extrapolation_List, Spatial_List=Spatial_List, Data_Geostat=alaska_pollock, PlotDir=DateFile )
```

```{r print_results, results="asis"}
pander::pandoc.table( Opt$diagnostics[,c('Param','Lower','MLE','Upper','final_gradient')] ) 
```

```{r diagnostics_encounter_prob, results="hide", eval=TRUE, tidy=TRUE, linewidth=50}
Enc_prob = SpatialDeltaGLMM::Check_encounter_prob( Report=Report, Data_Geostat=alaska_pollock, DirName=DateFile)
```


We can visualize fit to residuals of catch-rates given encounters using a Q-Q plot.  A good Q-Q plot will have residuals along the one-to-one line.  
```{r plot_QQ, eval=FALSE, tidy=TRUE, linewidth=50, message=FALSE, warning=FALSE}
#TmbData=TmbData; Report=Report; FileName_PP=paste0(DateFile,"Posterior_Predictive.jpg"); FileName_Phist=paste0(DateFile,"Posterior_Predictive-Histogram.jpg"); FileName_QQ=paste0(DateFile,"Q-Q_plot.jpg"); FileName_Qhist=paste0(DateFile,"Q-Q_hist.jpg")
Q = SpatialDeltaGLMM::QQ_Fn( TmbData=TmbData, Report=Report, FileName_PP=paste0(DateFile,"Posterior_Predictive.jpg"), FileName_Phist=paste0(DateFile,"Posterior_Predictive-Histogram.jpg"), FileName_QQ=paste0(DateFile,"Q-Q_plot.jpg"), FileName_Qhist=paste0(DateFile,"Q-Q_hist.jpg")) # SpatialDeltaGLMM::
```


```{r plot_years}
# Get region-specific settings for plots
MapDetails_List = SpatialDeltaGLMM::MapDetails_Fn( "Region"=Region, "NN_Extrap"=Spatial_List$PolygonList$NN_Extrap, "Extrapolation_List"=Extrapolation_List )
# Decide which years to plot                                                   
Year_Set = seq(min(alaska_pollock[,'Year']),max(alaska_pollock[,'Year']))
Years2Include = which( Year_Set %in% sort(unique(alaska_pollock[,'Year'])))
```

We can visualize many types of output from the model.  Here I only show predicted density, but other options are obtained via other integers passed to `plot_set` as described in `?PlotResultsOnMap_Fn`

```{r plot_density, message=FALSE, warning=FALSE, tidy=TRUE, linewidth=50}
Dens_xt <- SpatialDeltaGLMM::PlotResultsOnMap_Fn(plot_set=c(3), MappingDetails=MapDetails_List[["MappingDetails"]], Report=Report, Sdreport=Opt$SD, PlotDF=MapDetails_List[["PlotDF"]], MapSizeRatio=MapDetails_List[["MapSizeRatio"]], Xlim=MapDetails_List[["Xlim"]], Ylim=MapDetails_List[["Ylim"]], FileName=,,, Year_Set=Year_Set, Years2Include=Years2Include, Rotate=MapDetails_List[["Rotate"]], Cex=MapDetails_List[["Cex"]], Legend=MapDetails_List[["Legend"]], zone=MapDetails_List[["Zone"]], mar=c(0,0,2,0), oma=c(3.5,3.5,0,0), cex=1.8, plot_legend_fig=FALSE)
```

We can also extract density predictions at different locations, for use or plotting in other software. This is output in UTM using zone `r Extrapolation_List$zone-ifelse(Extrapolation_List$flip_around_dateline,30,0)`
```{r calc_density_dataframe, message=FALSE, warning=FALSE, tidy=TRUE, linewidth=50}
Dens_DF = cbind( "Density"=as.vector(Dens_xt), "Year"=Year_Set[col(Dens_xt)], "E_km"=Spatial_List$MeshList$loc_x[row(Dens_xt),'E_km'], "N_km"=Spatial_List$MeshList$loc_x[row(Dens_xt),'N_km'] )

Dens_DF %>% 
  as_data_frame() %>% 
  group_by(Year) %>% 
  summarise(ns = length(Density))
```

Ok, this thing clearly works, now I want to figure out how to get the actual predicted densities in space. 

# How to actually extract predicted densities?

`Report` is clearly the object of interest in here. Now, what the hell is in here. 

I think d_xcy is what you're looking for here, where for some unholt reason I think x is lon, y is lat, and c is some monstrosity from the depths of hell (category). OK no, y is year, since there are 35 years. So what in the holy hell is x? Knot?




# Try with Krigging......

from [here](https://rpubs.com/nabilabd/118172)

```{r}
View(alaska_pollock)
library(sp)
library(gstat)

alaska_pollock %>% as.data.frame %>% 
  ggplot(aes(Lat, Lon)) + geom_point(aes(size=Catch_KG), color="blue", alpha=3/4) + 
  ggtitle("Zinc Concentration (ppm)") + coord_equal() + theme_bw()

sp_alaska_pollock <-  alaska_pollock

coordinates(sp_alaska_pollock) <- ~ Lat + Lon


bbox(sp_alaska_pollock)

pollock.vgm <- variogram(log(Catch_KG + 1e-6)~1, sp_alaska_pollock) # calculates sample variogram values 

pollock.fit <- fit.variogram(pollock.vgm, model=vgm('Mat',300, 4.5)) # fit model


plot(pollock.vgm, pollock.fit) # plot the sample values, along with the fit model

extrap_points <-
  expand.grid(
  Lat = seq(
  min(alaska_pollock$Lat),
  max(alaska_pollock$Lat),
  length.out = 10
  ),
  Lon = seq(
  min(alaska_pollock$Lon),
  max(alaska_pollock$Lon),
  length.out = 10
  )
  )


plot1 <- sp_alaska_pollock %>% as.data.frame %>%
  ggplot(aes(Lon, Lat)) + geom_point(size=1) + coord_equal() + 
  ggtitle("Points with measurements")

# this is clearly gridded over the region of interest
plot2 <- extrap_points %>% as.data.frame %>%
  ggplot(aes(Lon, Lat)) + geom_point(size=1) + coord_equal() + 
  ggtitle("Points at which to estimate")

gridExtra::grid.arrange(plot1, plot2, ncol = 2)

coordinates(extrap_points) <- ~ Lat + Lon # step 3 above

pollock.kriged <- krige(log(Catch_KG + 1e-6) ~ 1, sp_alaska_pollock, sp_alaska_pollock, model=pollock.fit)

pollock.kriged %>% as.data.frame %>%
  ggplot(aes(x=Lat, y=Lon)) + geom_tile(aes(fill=var1.pred)) + coord_equal() +
  scale_fill_gradient(low = "yellow", high="red") +
  scale_x_continuous(labels=comma) + scale_y_continuous(labels=comma) +
  theme_bw()

```







